### 2017-01-17 Manuscripts/model-descriptions/copy-and-tweak/copy-and-tweak.tex

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{setspace, amsmath}

\begin{document}

\section{The Exemplar Copy-And-Tweak Model}

This model assumed that people represent categories as a collection of stored observations. When prompted to generate new examples, they copy and randomly tweak one of the stored examples. Thus, the model's generation is a two-part process:

\begin{enumerate}
  \item Select a source example from memory.
  \item Generate an example that is similar-but-not-too-similar to the source.
\end{enumerate}

More formally, let $x$ be a $j$-exemplar by $k$-feature matrix, corresponding to every stored exemplar. The probability that a source example $z$ is selected is given by:

\begin{equation}
  p(z) = exp(f_j) / \sum_j{ exp(f_j) }
\end{equation}

$f_j$ is a vector of values specifying each item's relative chance of being selected. In theory, $f_j$ may be set arbitrarily, but in practice it is sensible to differentiate between members of the target and contrast categories. For example, $f_j = -\gamma$ for members of contrast categories, and $f_j = \gamma$ for members of the target category. The resulting free parameter $\gamma$ ($>0$) thus controls the degree favoritism for members of the target category, with larger values producing larger differentials.

After a source exemplar $z$ is selected, similarity between candidate generation options $y$ is computed:

\begin{equation}
  s(y,z) = exp( -c \sum_k{|y_k - z_k|w_k})
\end{equation}

Where $c$ and $w_k$ are the standard specificity and attention weights discussed elsewhere. The probability a candidate stimulus will be generated is given by:

\begin{equation}
  p(y|z) \propto  
  \begin{cases}
     exp(\theta s(y,z)), & \text{if } s(y,z) \leq \phi \\
    0, & \text{otherwise}
  \end{cases}
  \label{eq:generation-choice}
\end{equation}

where $\theta$ is a response determinism parameter. $\phi$ is a similarity tolerance parameter ($0 \leq \phi \leq 1$) specifying the amount of dissimilarity tolerated between $y$ and $z$. Larger values allow for more similar examples to be generated. The final choice is made using a normalization of all possible values of $p(y|z)$.

\subsection*{The Practical Implementation}

In practice, however, it is useful to be able to generate model predictions more deterministically (i.e., without respect to any given source exemplar $z$). Generalizing things a little, the overall probability that an item $y$ will be generated is the sum of its generation probabilities given every possible source example $z$, $p(y|z)$, weighted by the probability each source will be selected, $p(z)$:

\begin{equation}
  p(y) = \sum_z{p(z)p(y|z) }
\end{equation}

In this way, source-independent probabilities can be obtained without monte-carlo simulation. First, the probability of each source $p(z)$ is computed (which is also required for simulation). Then, the pairwise similarity $s(y,z)$ between each source and each generation candidate is computed. Then the generation probability $p(y|z)$ is computed as per Equation \ref{eq:generation-choice}.  Then $p(y)$ is computed as above.

\end{document}