### 2017-01-30 Manuscripts/model-descriptions/jern-kemp-conjugate/jern-kemp-conjugate.tex
distance analyses, documentation update

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath,fancyvrb}

\begin{document}
\VerbatimFootnotes

\section*{Updated Version of the Jern \& Kemp (2013) Model}

Working backwards, what is needed for simulation of exemplar generation is the probability of generating a stimulus $y$ given exposure to members of the target category $x_B$:
\
\begin{equation}
p(y | x_B) = ?
\end{equation}
\
where $x_B$ may be empty. Jern \& Kemp's model achieves this using a generative process. Category members ($x_A$ or $x_B$; more generally written as $x_C$) are assumed to have been generated using an underlying category distribution (specifically, multivariate normal):
\
\begin{equation}
  x_C \sim Normal(\mu_{C}, \Sigma_{C})
\end{equation}
\
$p(y | x_C)$ is proportional to the candidate's density under the target category's distribution. Thus, obtaining the category distribution parameters $(\mu_{C}, \Sigma_{C})$ is key for generation. This document describes how we compute these variables in a conjugate model.

\subsection*{Computing $\mu_C$}
Assuming $(\mu_C, \Sigma_C)$ are Normal-Inverse-Wishart distributed (unknown mean, unknown variance):
\
\begin{equation}
  \mu_C = \dfrac
    {\kappa\mu_{0} + n_C \bar{x_C}}
    {\kappa + n_C}
    \label{eq:category_mus}
\end{equation}
\ 
where:
\begin{itemize}
    \setlength\itemsep{-0.5em}
    \item $\mu_{0}$ is the prior mean along $p$ dimensions. Here we set it to the middle of the space.
    \item $\kappa$ is a scalar hyper-parameter, roughly weighting the importance of $\mu_{0}$. $\kappa$ must be greater than zero.
    \item $n_C$ is the number of observations in $x_C$
    \item $\bar{x_C}$ is the sample mean along $p$ dimensions
\end{itemize}

In the case of a populated class, $\mu_{C}$ ends up lying somewhere between $\mu_{0}$ and $\bar{x_C}$, depending on $\kappa_{0}$ and $n_C$. In the case of an empty class, $n_C = 0$, Equation \ref{eq:category_mus} reduces to $\mu_{C} = \mu_{0}$. Because we set $\mu_0$ to the center of the space, this outcome is the same as if we had integrated over all possible $\mu_C$. In practice, if $n_C = 0$, the model picks a stimulus at random from all candidates (uniform probabilities).

\subsection*{Computing $\Sigma_D$}

Unlike $\mu_C$, $\Sigma_C$ cannot be computed considering only the members of category $y$. Instead, $\Sigma_C$ is influenced both by the distribution of $x_C$ and by members of other categories through $\Sigma_D$.

$\Sigma_D$ is inferred based on the observed (empirical) category covariances $C_y$. We assume these covariances to be Wishart-distributed, and so $\Sigma_D$ can be computed as:

\
\begin{equation}
    \Sigma_D = \Sigma_0 + \sum_{C}{C_C}
\end{equation}
\
$\Sigma_{0}$ is a $p$-by-$p$ prior covariance matrix. We use a $p$-dimensional identity matrix $I_p$ multiplied element-wise against a free parameter, $\rho$, controlling the amount of variance assumed by the prior:
\
\begin{equation}
    \Sigma_0 =  I_p\rho
\end{equation}

Thus, categories are assumed to have some degree of variance along each feature (specified by $\rho$), but not are assumed to possess feature-feature correlations.

\subsection*{Computing $\Sigma_C$}

Assuming $(\mu_C, \Sigma_C)$ are Normal-Inverse-Wishart distributed, $\Sigma_C$ can be computed as:

\begin{equation}
  \Sigma_C = [\Sigma_D\nu + C_C +
    \dfrac
    {\kappa n_C}
    {\kappa + n_C}
    (\bar{x_C}-\mu_C)(\bar{x_C}-\mu_C)^T
  ] (\nu + n_C)^{-1}
  \label{eq:Sigma_y}
\end{equation}
\
$\kappa$, $\bar{x_C}$, $C_C$, $n_C$, $\mu_0$,  are the same values as described above. $\nu$ is an additional free parameter, weighting the importance of $\Sigma_{D}$. $\nu$ must be greater than $p-1$. When $x_b$ is empty, Equation \ref{eq:Sigma_y} reduces to $\Sigma_C = \Sigma_D$.

\subsection*{Computing response probabilities $p(y | x_C)$}

$x$ are assumed to be drawn from a the distribution given by $Normal(\mu_{C}, \Sigma_{C})$. Thus,

\begin{equation}
  p(y) \propto Normal(y; \mu_{C}, \Sigma_{C})
\end{equation}

In practice, $p(y)$ is computed by first obtaining the relative density of every possible generation candidate $y_i$ under the category distribution. The end probability is a normalization of these values:

\begin{equation}
  p(x) = \dfrac
    {exp( \theta Normal(y ; \mu_{C}, \Sigma_{C}))}
    {\sum_i exp( \theta Normal(y_i ; \mu_{C}, \Sigma_{C}))} 
\end{equation}
\
where $\theta$ is a response determinism parameter.

\subsection*{Description of free parameters}

\begin{itemize}
    \setlength\itemsep{-0.5em}
    \item $\kappa$. Scalar, $\kappa>0$. Weights the importance of $\mu_{0}$ in inferring category $\mu_{C}$.
    \item $\rho$. Scalar, $\rho>0$. Sets the assumed variance in the domain prior, $\Sigma_{0}$.
    \item $\nu$. Scalar, $\nu > p-1$. Weights the importance of $\Sigma_{D}$ in inferring the domain $\Sigma_{C}$.
    \item $\theta$. Scalar, $\theta > 0$. Response determinism parameter.
\end{itemize}

\end{document}