### 2017-06-09 Manuscripts/cog-psych/cover-letter.tex
cover letter drafted; really tremendous work, outstanding

%!TEX output_directory = latex_out/

\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

We are pleased to submit to {\em Cognitive Psychology} a manuscript entitled ``Creating Something Different: Similarity and Contrast in Concept Generation.''

The manuscript reports two experiments and formal modeling on the topic of category generation. Previous work on the topic has not considered the integral role of contrast in ensuring generated categories differ from what is already known. Our behavioral experiments demonstrate such an influence, and our modeling efforts show that a novel exemplar-based approach (based on the Generalized Context Model) can explain the influence of category contrast in the generation process.

We expect that a broad range of readers will find our results noteworthy and compelling.  The work reported in the manuscript has not been submitted for journal publication. A previous version of this work is to appear in the Proceedings of the Thirty-Ninth Annual Conference of the Cognitive Science Society (Conaway \& Austerweil, 2017).

Thank you for considering our work for publication. Correspondence regarding the manuscript may be addressed to:

Joseph L. Austerweil \\
1202 West Johnson Street,\\ 
Madison, WI 53706.\\
E-mail: austerweil@wisc.edu

\end{document}