### 2018-04-06 Presentations/mathpsych_2018/main.tex

%%
\documentclass[man,12pt]{apa6}
\title{Learning generated categories: Applying PACKER to category assignments}
\shorttitle{PACKER in category learning}
\fourauthors{Shi Xian Liew}{Nolan Conaway}{Kenneth Kurtz}{Joseph Austerweil}
\fouraffiliations{University of Wisconsin -- Madison}{Shutterstock}{Binghamton University}{University of Wisconsin -- Madison}

%Turn off ligatures for easy copypasta
\usepackage{fontspec}
\setmainfont[Ligatures={NoRequired,NoCommon,NoContextual}]{Times New Roman}

\leftheader{Beitzel}

\abstract{\input{abstract}}

%\keywords{APA style, demonstration}

\begin{document}
\maketitle
\end{document}