### 2018-04-20 Manuscripts/model-descriptions/packer/gamma_invariance.tex
Trying out another version of PACKER, where this time tradeoff is implemented as weights for both contrast and target similarities.

\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
headheight=15pt]{geometry} \usepackage{amsmath, listings}

\begin{document}

\section{Description of the non-variant effect of gamma on generation and
assignment probabilities in PACKER}

This document refers to the version of PACKER as described in the manuscript
submitted to Cognitive Psychology (sometime in 2017; henceforth referred to as
the manuscript). In this version, PACKER performs very well in predicting the
generation of new category exemplars. In particular, the tradeoff parameter
(henceforth $\gamma$) actually affects the predictions generated in an expected
way. However, when trying to produce predictions of category assignments,
varying $\gamma$ does not affect the predictions in any way. The following
section of this document describes why this is occurring.

Equation 2 in the manuscript specifies that the aggregated similarity $a$
between generation candidate $y_i$ and a vector of stored exemplars $x$ can be
expressed as:

\begin{equation} a(y_i, x) = \sum_j{f(x_j) s(y_i, x_j)}
\label{eq:aggsim}
\end{equation} where $f(x_j) = \gamma$ when $x_j$ is a member of a target
category $T$ and $f(x_j) = \gamma - 1$ when $x_j$ is a member of a contrast
category $C$.

We can express Equation \ref{eq:aggsim} here a little differently:

\begin{equation} a(y_i,x) = \gamma \sum_{k \in K_T}{s(y_i,x_k)} + (\gamma-1)
\sum_{k \in K_C}{s(y_i,x_k)}
\label{eq:aggsimsplit}
\end{equation} where $K_T$ is the set of all target category members and $K_C$ is
the set of all members of the contrast category. To make the algebra a little
less tedious we can simplify some of the representations here. Let $t_i =
\sum_{k \in K_T}{s(y_i,x_k)}$ and $c_i = \sum_{k \in K_C}{s(y_i,x_k)}$, which allows
us to express Equation \ref{eq:aggsimsplit} as:

\begin{equation} a(y_i,x) = \gamma t_i + (\gamma-1) c_i .
\label{eq:aggsimsimp}
\end{equation}

According to the implementation of PACKER in the manuscript, an exponentiated
Luce's choice rule is used to find the probability $p$ of generating candidate
$y_i$:

\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot a(y_i, x)\}} {\sum_i{\exp
\{\theta \cdot a(y_i,x)\}}} .
\label{expluce}
\end{equation} Thus far, Equation \ref{expluce} only apply to predictions of
category generation. In order to extend PACKER to category assignment
predictions, we need to find not the probability of generating candidate $y_i$,
but rather, the probability of assigning candidate $y_i$ to the target category. 
 these equations only apply to predictions of category
generation. However, predictions of category assignments require some
reconceptualization of these equations, although most of the equations retain
the same structure. Specifically, for Equations \ref{eq:aggsim} to
\ref{eq:expluce}, $i$ now iterates over each category (as opposed to each
generation candidate; i.e., $i \in K$ $ y_i$ now represents the assignment of candidate $y$ to
target category $i$.

Substituting $a(y_i,x)$ with Equation \ref{eq:aggsimsimp}:
\begin{equation} p(y_i) = \dfrac{\exp \{\theta \cdot \gamma t_i + (\gamma-1)
c_i\}} {\sum_i {\exp \{\theta \cdot \gamma t_i + (\gamma-1) c_i\}}}
\label{eq:expluceexp}
\end{equation}

Whereas the denominator in Equation \ref{eq:expluceexp} is the summation of
aggregated similarities over all generation candidates in category generation,
category assignment predictions require that the denominator specifies the
summation over all category assignments. Specifically, $i$ in the summation term
represents each category, as opposed to each generation candidate.

This model assumes that people represent categories as exemplars in a
multidimensional space, and that generation is constrained by both similarity to
members of the category being generated \textit{and} dissimilarity to members of
other categories. The model assumes people generate categories that are
dissimilar to known categories and have strong within-class similarity.

The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's
law:

\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
\label{eq:similarity}
\end{equation}

When prompted to make a generation decision, participants are thought to
consider both similarity to examples from other categories as well as similarity
to examples in the target category. More formally, the aggregated similarity $a$
between candidate $y$ and the model's stored exemplars $x$ can be computed as:

\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
\end{equation}

Where $f(x_j)$ is a function specifying each stored example's degree of
contribution toward generation. Although $f(x_j)$ may be set arbitrarily, in
PACKER it is set according to class assignment. For known members of the target
category, $f(x_j) = \gamma$. For members of contrast categories, $f(x_j) =
\gamma - 1$. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$)
controlling the trade-off between within-class similarity and between-class
dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category
members, and $\gamma = 0$ produces exclusive consideration of opposite-category
members. When $\gamma = 0.5$, the similarity to contrast categories is
effectively subtracted from the similarity to the target category.

The probability that a given item $y$ will be generated given the model's memory
$x$ is computed using relative summed similarity values across all generation
candidates $y_i$:

\begin{equation} p(y) = \dfrac { \exp \left \{ \theta \cdot a \left( y, x
\right) \right \} } { \sum_i{ \exp \left \{ \theta \cdot a \left( y_i, x \right)
\right\} } }
\label{eq:packer-choice}
\end{equation}

Where $\theta$ ($\geq 0$) is a free parameter controlling overall response
determinism.

\end{document}