### 2018-05-24 Manuscripts/cog-psych/revision/methods_results_e3.tex

% \documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
% headheight=15pt]{geometry} \usepackage{amsmath, listings}

% \begin{document}

\section{Participants, Materials, and Procedure}

Experiment 2 recruited 122 participants who each generated one set of four Alpha
and four Beta category exemplars. Among these 122 generated category sets were
102 unique sets (i.e., they contained a unique collection of Alpha and Beta
exemplars). Consequently, for Experiment 3, we recruited 102 participants with
one participant presented with a different unique category set.

Participants observed four blocks of eight trials. Each trial began with the
presentation of a fixation cross for 500 ms. This was followed by the
presentation of one exemplar randomly sampled without replacement from the
unique category set. Participants were tasked with assigning the presented
exemplar to either the Alpha or Beta category with no time limit imposed.
Feedback was automatically displayed for 2500 ms after each response.

\section{Results}

Overall accuracy of the participants was high, with a mean error rate of $.19$
($SD = .19$).

[Place error curve plots here, not unlike what NGPMG1994 has done]

[Move on to modelling analyses. Include correlation plots, don't worry about
individual fits. Also include the varying of PACKER's performance (and maybe
representativeness?) on correlation with changing tradeoff. ]

[Note Joe's comments on figure formatting (font size, remove redundant y-axes, etc)]