### 2018-06-28 Manuscripts/model-descriptions/packer/packer_new.tex
Likelihood ratio tests for correlation might not be appropriate. Until I find a better solution, run bootstraps first.

%!TEX output_directory = latex_out/

\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
headheight=15pt]{geometry} \usepackage{amsmath, listings}

\begin{document}

\section{The PACKER Model: Producing Alike or Contrasting Knowledge with
Exemplar Representations}

This model assumes that people represent categories as exemplars in a
multidimensional space, and that generation is constrained by both similarity to
members of the category being generated \textit{and} dissimilarity to members of
other categories. The model assumes people generate categories that are
dissimilar to known categories and have strong within-class similarity.

The similarity between exemplars $x_i$ and $x_j$ is computed using Shepard's
law:

\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
\label{eq:similarity}
\end{equation}

When prompted to make a generation decision, participants are thought to
consider both similarity to examples from other categories as well as similarity
to examples in the target category. More formally, the aggregated similarity $a$
between candidate $y$ and the model's stored exemplars $x$ can be computed as:

\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
\end{equation}

Where $f(x_j)$ is a function specifying each stored example's degree of
contribution toward generation. Although $f(x_j)$ may be set arbitrarily, in
PACKER it is set according to class assignment. For known members of the target
category, $f(x_j) = \gamma$. For members of contrast categories, $f(x_j) =
\gamma - 1$. $\gamma$ is thus a free parameter ($0 \leq \gamma \leq 1$)
controlling the trade-off between within-class similarity and between-class
dissimilarity: $\gamma = 1$ produces exclusive consideration of same-category
members, and $\gamma = 0$ produces exclusive consideration of opposite-category
members. When $\gamma = 0.5$, the similarity to contrast categories is
effectively subtracted from the similarity to the target category.

The probability that a given item $y$ will be generated given the model's memory
$x$ is computed using relative summed similarity values across all generation
candidates $y_i$:

\begin{equation} p(y) = \dfrac { \exp \left \{ \theta \cdot a \left( y, x
\right) \right \} } { \sum_i{ \exp \left \{ \theta \cdot a \left( y_i, x \right)
\right\} } }
\label{eq:packer-choice}
\end{equation}

Where $\theta$ ($\geq 0$) is a free parameter controlling overall response
determinism.

\end{document}