### 2018-07-17 Manuscripts/cog-psych/revision/cogpsych.tex
Added representativeness to manuscript. Updated figures and results. Discussion in various places and some reviewer-related responses still need work.

%!TEX output_directory = latex_out/
\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{csquotes}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage[toc]{appendix}
\usepackage[natbibapa]{apacite}

% set up PGF
\pgfplotsset{compat=1.6}
\newcommand\inputpgf[2]{{
\let\pgfimageWithoutPath\pgfimage
\renewcommand{\pgfimage}[2][]{\pgfimageWithoutPath[##1]{#1/##2}}
\input{#1/#2}
}}

% set up notes-- different backgrounds for Nolan and Joe!
\newcommand\nbcnote[1]{\todo[inline, backgroundcolor = yellow]{\textbf{NBC}: #1}}
\newcommand\jlanote[1]{\todo[inline, backgroundcolor = lime]{\textbf{JLA}: #1}}
\newcommand\sxlnote[1]{\todo[inline, backgroundcolor = cyan]{\textbf{SXL}: #1}}

% set up header % 
\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\lhead{RUNNING HEAD: Similarity, Contrast, and Representativeness in Categorization}
\fancyhead[R]{\thepage}

\begin{document}

% ------- TITLE PAGE ------- %
\begin{center}
\hfill
\\[1in]

Creating and Learning Something Different: \\
Similarity, Contrast, and Representativeness in Categorization

\vfill
Joseph L. Austerweil\textsuperscript{1}
Nolan Conaway\textsuperscript{1}, 
Shi Xian Liew\textsuperscript{1},
and Kenneth J. Kurtz\textsuperscript{2}
\\[\baselineskip]
\textsuperscript{1}Department of Psychology, University of Wisconsin-Madison, Madison, WI, USA
\textsuperscript{2}Department of Psychology, Binghamton University, Binghamton, NY, USA
\\[1in]

\vfill

Author Note

Correspondence concerning this article should be addressed to: 
Joseph L. Austerweil, 1202 West Johnson Street, Madison, WI 53706.
E-mail: austerweil@wisc.edu

\end{center}
\clearpage

% ------- ABSTRACT PAGE ------- %
\doublespacing
\section*{Abstract}
\jlanote{need to update} 
\sxlnote{updated} 
%
The ability to creatively generate new concepts and ideas is among the most
fascinating aspects of human cognition, but we do not have a strong
understanding of the cognitive processes and representations underlying concept
generation. In this paper, we study the generation of new categories using the
computational and behavioral toolkit of traditional artificial category
learning. Previous work in this domain has focused on how the statistical
structure of known categories generalizes to generated categories, overlooking
the degree of contrast between the known and generated categories as a factor.
We report three experiments demonstrating that contrast between what is known
and what is created is of fundamental importance for categorization and the
creative process. We propose two novel approaches to modeling category contrast:
one focused on exemplar dissimilarity and another on the representativeness
heuristic. These new models are compared to two key alternatives. Our
experiments and simulations demonstrate specifically how category contrast
influences creative generation, and how it can interact with other constraints
to produce different types of creations. Our work also serves as a concrete
example of how well-established and highly controlled experimental frameworks
(such as those in the field of category learning) can be used to make progress
in an intriguing yet understudied domain.

\setlength\parindent{0.5in}
{\em Keywords}: categorization; concepts; creativity; generation; computational modeling; exemplar models
\clearpage

% % OLD OLD INTRO:
% Creativity, innovation, imagination, and the creation of new ideas are among the most fascinating, important and difficult human capabilities to study. They are fascinating and important capabilities because they have produced some of the greatest scientific breakthroughs that revolutionized the world. Unfortunately, technological and scientific breakthroughs are rare and the result of a great deal of cognitive effort, which makes it difficult to investigate the underlying cognitive representations and processes responsible for them using standard experimental methodology or to formalize them in computational models. Although we tend to focus on the most salient products of these processes (e.g., scientific breakthroughs), every person is likely to have generated many novel sentences, thoughts, and/or drawings given the combinatorial explosion of possible statements given a reasonably-sized set of primitive elements and combination rules in language, thought, and perception \citep{goldstone2003}. This observation provides a fantastic opportunity: By examining how people create new concepts in a domain amenable to formalizing in computational models, we can investigate (some aspects of) creativity, innovation, and imagination scientifically.

% Towards the aim of providing a formal account of the cognitive mechanisms involved in creativity and innovation, we explore a form of creative cognition that can be studied using standard behavioral methodology and is amenable to formal modeling: category generation. In a category generation task, participants are given some background knowledge about a domain (a cover story and/or learning one or more categories in the domain) and then are asked to generate another category in the domain. Unlike creativity and innovation, categorization is well studied from a breadth of perspectives, including behavioral, neural, comparative, and computational  \citep{kurtz2015human,mack2013,margolis2015,pothoswills2011}. By analyzing a creative task using formal tools from the categorization literature, we can formalize some theories of creativity and test them empirically in a more rigorous manner.

%cut from submission intro:

%Yet, it is also one of the most difficult human capabilities to study.  We tend to focus on the most salient products of conceptual generation (e.g., scientific breakthroughs), but every person is likely to have generated many novel concepts in ordinary life. Indeed, the generation of novel sentences, works of art, and so on may also be considered products of our capacity to create new concepts. However, while this capacity is of great interest, it is also highly complex, making it difficult to study in a typical laboratory setting. By consequence, we do not have a strong understanding of the representations and processes involved in the generation of new concepts.

%Questions concerning the generation of novel concepts are typically addressed within research on creativity, a tremendously diverse field employing a multitude of experimental and theoretical approaches \citep[see][]{kaufman2010handbook}. The processes underlying conceptual generation are commonly investigated using the creative cognition approach \citep{finke1992creative, smith1995creative}, in which the products of creative acts (such as drawings) are analyzed to obtain insights into how they might have been created. These approaches have made great advances in our understanding of creativity, but the focus on particular types of creative products such as drawings makes it difficult to formalize the proposed processes and representations within a computational modeling framework. As a consequence, this approach does not often employ computational models to provide formal tests of competing theories.

% ------- BEGIN! ------- %
\begin{flushleft}

\section{Introduction}
\setlength\parindent{0.5in}

Creating and learning new ideas is one of the most fascinating and important
human capabilities. For example, cell phones and smart phones are newly created
categories of objects that young and old adults learned and have become reliant
on in the last few decades. In fact, computational researchers have identified
generating novel objects and objectives as one of the defining, and most
difficult to formalize, characteristics of intelligent life
\citep{lake2017,lehman2011,taylor2016}. Yet, the cognitive mechanisms that
enable people to innovate and learn new innovations are not well understood and
are understudied by scientists. In part, this is due to the inherent difficulty
of designing and conducting experiments that test these capabilities. How would
a scientist devise an experiment that induces a participant to create and learn
new categories that are as interesting as cell phones and smart phones? In this
article, we use computational models and behavioral experiments to investigate
the prior expectations people have when they learn and generate new categories.

Generating and learning new concepts are, however, not altogether different from
the types of behaviors typically studied in cognitive psychology laboratories.
In particular, generating a member of a novel class can be considered a `special
case' application of existing category knowledge
\citep{kemp2014taxonomy,kurtz2015human}. Research in categorization typically
focuses on what properties of a category affect the classification of an
object's category given its features \citep{shepard1961learning,kurtz2013human},
and the prediction of an object's unobserved features given some of its other
features and/or its category \citep{markman2003category}. The generation of
members of a new category consists of inferring {\em all} features for a {\em
novel} category label. Thus, we can make progress formalizing the processes
involved in category generation by extending theories of categorization to this
case. Then, to ensure that these principles reflect basic principles of
categorization and are not just specific to category generation, we can examine
what factors of the generated categories affect the ease at which people learn
the category.

Previous work in category learning and generation has established that people
are highly sensitive to the structural properties of categories, such as
correlations between the features of category members and the relation between
items within the same category and those in different categories
\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
Inspired by this work, previous research on the topic of category generation has
explored a similar principle: People tend to create new categories that have
similar {\em statistical regularities} as previously learned categories
\citep{jern2013probabilistic,ward1994structured}. Although this is an important
characteristic of generating new categories, it cannot be the only one. Taken to
the extreme, the best ``new'' category in terms of having the same statistical
regularities to other categories would be identical to a known category that is
representative of the domain (and thus, not new at all).

\sxlnote{I think we can address Charles' point on real world examples of how
  contrast can be important around here.}
To successfully generate something novel, what is generated must be different
from what is already known. This fundamental constraint, ``being different'', or
contrasting from other categories in the relevant domain, is the focus of our
work. Although implicitly assumed in some work, this constraint has been
overlooked in previous research: To our knowledge, there has not been any
systematic investigation addressing how generated categories {\em differ} from
what is already known. Although the idea of category contrast is ubiquitous
throughout the categorization literature, and extends to a variety of other
fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
be ``different'' is vague, as there are many ways it could be different from a
previously observed category.

We examine two different definitions of how categories should differ from one
another: exemplar dissimilarity or representativeness of the alternative
category. To formalize the first technique, we build on the largely successful
exemplar modeling framework that
\citep{medin1978context,nosofsky1984choice,nosofsky1986attention}. In doing so,
we propose a novel exemplar model of category generation, {\em Producing Alike
  and Contrasting Knowledge using Exemplar Representations} (PACKER),
formalizing how new categories should differ from previous categories. It
embodies the exemplar dissimilarity principle by incorporating a factor that
repulses members of opposite categories from one another. To foreshadow, we find
that this notion of contrast predicts human category learning performance, above
and beyond other category learning models in a novel behavioral experiment. This
demonstrates that contrast plays a central role in other important
categorization tasks, which is not specific to category generation.

The second hypothesis for how contrast might affect categorization is contrast
as {\em representativeness}. The representativeness heuristic \citep{kahneman73}
states that people make judgments regarding an outcome based on how
representative it is of the evidence given in the current context. The heuristic
is a powerful theoretical construct that has been used to capture a range of
complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
especially those that deviate from normative theory (as given by a
straightforward application of Bayes' rule). For example, the coinflip sequence
TTHTH is perceived to be more random (due to it being representative of a
coinflip sequence) than TTTTT (despite both having the same probability of
having been generated from a fair coin). Although there is a healthy debate as
to whether perceived randomness {\em really} deviates from normative theory
\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
main explanations of human judgments \citep{reimers18}. To formalize the
hypothesis that people expect a new category to be representative of the
opposite of the current category, or contrast as {\em representativeness}, we
use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
has been used successfully to capture human performance in color categorization
\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
learning \citep{rafferty10}. We do so by extending the state-of-the-art category
generation model \citep{jern2013probabilistic} and find that human category
generation is best captured as generating representative, rather than probable,
samples.

The outline of the article is as follows. First we describe previous empirical
work on the topic of category generation, as well as the computational
formalizations of the theories in those reports. We then present two hypotheses
for how contrast might affect categorization and formalize them in computational
models. The first is a novel computational model, which is designed to generate
categories that systematically differ from existing categories in the domain.
The second assumes the goal of category generation is to create representative
samples of the opposite of the observed categories. We present two experiments
demonstrating strong and systematic effects of category contrast on concept
generation, and we qualitatively and quantitatively analyze the performance of
each model in capturing human category generation. Afterwards, we find support
for category contrast predicting the ease at which people learn a pair of novel
categories in an empirical experiment. We conclude with a discussion of the
implications of our results for categorization and creative cognition, and
directions for future work.

\section{Prior work}

Much of what we know about concept generation comes from the foundational
literature on creative cognition. In a series of reports, Ward and colleagues
\citep{ward1995s,ward1994structured,marsh1999inadvertent,ward2002role,smith1993constraining}
established that category generation is highly constrained by prior knowledge:
Generated categories tend to consist of features observed in known categories,
and they tend to exhibit the distributional properties as found in known
categories. In a seminal study, \cite{ward1994structured} asked participants to
generate new species of alien animals by drawing and describing members of the
species. People tended to generate species with the same features as on Earth
(e.g., eyes, legs, wings), and possessing the same feature correlations as on
Earth (e.g., feathers co-occur with wings). Likewise, aliens drawn from the same
species tended to share more features with one another compared to members of
opposite species.

\sxlnote{Added thomas98 description here} It is worth pointing out that the
observation of regularities in distributional properties across categories is
not confined to the work of Ward and colleagues. Most notably, \citet{thomas98}
found that when people were asked to generate feature values for exemplars with
a missing feature (specifically circles of different sizes, each with a radial
line of varying orientation), the resulting set of exemplars tended to share the
same feature correlations as exemplars from previously learned categories. [the
transition to the next point isn't great, though]

The broader set of observations made by Ward and colleagues provide a great deal
of insight into the role of prior knowledge in constraining category generation.
Much of the work from this area
\citep[e.g.,][]{smith1993constraining,marsh1999inadvertent} focuses on how
information provided to participants (such as an example of a species generated
by other participants) can drastically diminish difference of a new category
from pre-existing categories. Theoretical accounts of these effects have
primarily been grounded within the categorization literature. For example, the
predominant ``Path of Least Resistance'' account
\citep[see][]{ward1994structured,ward1995s,ward2002role} proposes that, when
generating a new species of animal, people retrieve from memory a known
subcategory of animals (e.g., {\em bird}, {\em dog}, {\em horse}), and simply
change some of the features to make something new. People are thought to change
only features that are not characteristic of the retrieved category (e.g., if
{\em bird} was retrieved, the presence of {\em wings} would not change, but {\em
color} might). This theory incorporates elements of the highly influential
basic-level categories framework \citep{rosch1975cognitive,rosch1976basic}, as
well as the exemplar view \citep{medin1978context,brooks1978nonanalytic}. While
this work is been incredibly useful in providing a conceptual sketch of
generation theories, the hand-drawn responses used in the experiments paradigms
precludes the development of formal approaches that can be used to test the
theories in a fine-grained manner.

\cite{jern2013probabilistic} recently showed that concept generation could be
studied in a more controlled manner through the well-developed methods of an
artificial categorization paradigm \citep[see][for a review]{kurtz2015human}. In
Experiments 3 and 4 of their article, participants were exposed to members of
experimenter-defined categories of "crystals" varying in size, hue, and
saturation. Following a training phase during which the experimenter-defined
categories were learned, participants were asked to generate novel categories of
crystals. In a finding mirroring \cite{ward1994structured},
\cite{jern2013probabilistic} found that participants generated categories with
the same distributional properties as the experimenter-defined categories. For
example, after training on categories with a positive correlation between the
size and saturation features (larger sized crystals were more saturated),
participants generated novel categories with the same positive correlation. This
finding is notable, as it demonstrates that category generation can be studied
in a well-known and highly controlled experimental paradigm.

The authors evaluated the predictions of several formal models on their data.
Most notably, they showed that a hierarchical Bayesian model provided the
strongest account of their results. Their model views observed examples as
samples from an underlying category distribution, describing the location of the
category in the space, as well as how it varies along each feature. In turn,
each category is viewed as a sample from an underlying {\em domain}
distribution, specifying distributional commonalities among the observed
categories. Generated categories are thought to stem from the same domain
distribution as observed categories, thus the distributional properties of
observed categories will be preserved within the generated category.

\cite{jern2013probabilistic} additionally tested a ``copy-and-tweak'' model that
broadly resembles the earlier ``Path of Least Resistance'' account. The core
proposal is that participants generate new items by copying stored examples from
memory and tweaking them to generate something new. The copy-and-tweak model
differs from the Path of Least Resistance account in that it notably omits the
hierarchical organization of categories, as well as selectivity in which
features are changed \citep[both of which are factors in the Path of Least
Resistance account;][]{ward2002role}. Instead, their copy-and-tweak model
corresponds to a direct exemplar-similarity approach
\citep[e.g.,][]{nosofsky1984choice,nosofsky1986attention}, generating new items
according to their similarity to known members of the target category. The
copy-and-tweak model provided a poor account of their results, as the
experiments devised by \citeauthor{jern2013probabilistic} were specifically
designed to challenge it. However, its application is notable as a first step
toward understanding theories developed in the creative cognition literature
using well-known formal approaches from the categorization literature.

\section{And Now for Something Different: A Role For Contrast}

Prior work on category generation has explored only one factor: distributional
correspondences between learned and generated categories. As a result most of
the computational, theoretical, and empirical efforts have been focused on
explaining those effects. In this paper, we investigate another important
factor: category contrast. To generate a novel concept, individuals must produce
something that is in some capacity {\em different} from what they already know.
By consequence, we propose that contrast should be a primary constraint that
guides how people create new categories: New categories should be different from
existing ones.

\jlanote{check to make sure the next paragraph has been updated enough}
\sxlnote{I think it's updated enough! If given a chance I might edit this a
  little to be less harsh on hierarchical bayesian models in general though I
  think, since the representativeness model is one of them?}
Although it is evident that people are {\em capable} of creating new concepts
and categories, it is not entirely clear how new concepts are systematically
made different from what is already known. The hierarchical Bayesian model
developed by \cite{jern2013probabilistic} assumes that differences between
observed and generated categories are only due to random variation. The model
assumes that generated categories are sampled from the same underlying domain
distribution as observed categories, and will thus share a common distributional
structure. The model does not make predictions about the {\em location} of the
category within the domain (the perceptual instantiation of category members).
Under a strict interpretation of their model, given knowledge of a single
category within the domain, the most probable new category to be generated is
located in {\em exactly} the same location and possesses an identical
distributional structure. This is not an issue with their model specifically,
but with a broader class of standard hierarchical Bayesian models \citep[e.g.,
][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
the latent generative process the same underlying distribution generates all of
the categories and thus, any differences between categories are due to {\em
noise} and should not be {\em systematic}. If the goal of creating a new
category given others in a domain is generating from the posterior predictive
distribution then the best a standard hierarchical Bayesian model without a
notion of contrast built into the model can do at capturing contrast is to
assume that the new category is placed uniformly at random over stimulus space.
But, this defeats the purpose of a hierarchy as it is ignored when determining a
new category location!\footnote{It is plausible that some hierarchical Bayesian
model could be created that generates categories different from each other.
However, this model would not be a standard application or extension of most
pre-existing hierarchical Bayesian models. The generative process would need to
include a component that promotes contrast.} Note that this does not mean
location information and contrast cannot be captured using hierarchical Bayesian
models. In fact, our model of contrast as representativeness is a hierarchical
Bayesian model, but with a different goal than predicting the next category.

The copy-and-tweak model tested by \cite{jern2013probabilistic} also claims
little about how generated categories should contrast with what is already
known. In their simulations, the model was only tested on generation after the
learner had been exposed to members of the target category, and so the model's
ability to generate a new category from scratch was not evaluated. However, the
model's generation is based exclusively on similarity to known members of the
{\em target} category; when there are no members of the target category,
generation is presumably random.

\sxlnote{Charles suggested we either remove the section below, or move it to the
  discussion. I've tried finding a good place for it in the discussion but it
  doesn't fit very well, and moving it takes away this nice motivation for
  our further investigation into contrast. I think that keeping it here in the
  introduction might work out, provided Charles finds my argument for why the
  regularity assumption is too strong convincing enough. More details a little
  further down.}
\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}

Although existing accounts of category generation broadly overlook the role of
category contrast in determining what is novel versus familiar, it was
implicitly assumed that learners in previous experiments were {\em successful}
in creating new categories. Thus, if contrast plays a robust role in category
genration, its effects should be discernable within the experimental results of
these studies. To provide a test of the influence of category contrast within
existing data, we conducted a novel analysis of Experiment 3 from
\cite{jern2013probabilistic}.\footnote{Although \cite{jern2013probabilistic}
reported four experiments, we focus on Experiment 3 because their first two
experiments tested generation of items into known categories, and their fourth
experiment was identical to Experiment 3 but with a restricted generation
space.}

Participants in their experiment were exposed to members of two
experimenter-defined categories of `crystals' varying in hue, saturation, and
size. Each category possessed a unique hue, but varied in saturation and size:
In the `Positive' condition, there was a positive correlation between these
features (i.e., larger sized crystals were more saturated), and in the
`Negative' condition this relation was reversed. In the `Neutral' condition,
there was no correlation between saturation and size. After learning about the
categories from each condition, participants were asked to generate six
exemplars belonging to a novel class. As noted above,
\cite{jern2013probabilistic} found that the generated categories tended to
follow the distributional properties of the experimenter-defined categories:
Generated categories were tightly distributed along the hue feature, and
possessed the same saturation-size correlations as in the learned categories.
\cite{jern2013probabilistic}, however, did not analyze or discuss how the
generated categories {\em differed} from the experimenter-defined categories.

Because each experimenter-defined category in the \cite{jern2013probabilistic}
experiment possessed a distinct hue shared by all members of the category, it is
sensible that participants might generate a category with a hue distinct from
the experimenter-provided categories. If creative generation were influenced by
category contrast in this way, the hues of generated categories should be
systematically different from those of the experimenter-defined categories.
Unfortunately, stimulus hue was encoded and presented in the
Hue-Saturation-Value (HSV) color space, which is device-dependent and not
perceptually normed such that perceived color similarity corresponds to
proximity in the color space (as opposed to a color space such as CIELAB that is
device-independent and equidistant sets of points correspond to pairs of colors
that have the same perceptual similarity; \citealp{wyszecki1967}). Further, they
did not calibrate their monitor, and so we cannot know the precise colors
presented to participants. As \cite{jern2013probabilistic} were interested in
relations between the saturation and length of examples in generated into novel
categories, these issues do not undercut their analyses and results. However,
these issues pose a significant challenge to evaluating contrast between the
experimenter-defined and participant-generated categories along the hue
dimension. It is plausible that two uncalibrated monitors could display the same
HSV color and the colors be perceived in different color categories (especially
for color boundaries that vary over lightness, such as the yellow-brown
boundary).

Although we cannot know the precise colors that were displayed or perceived, we
can still analyze their results from a coarse perspective to see whether there
is preliminary support for contrast. To do so, we binned all possible hues into
one of eight uniformly-spaced color groups: \{{\em Red}: $0-0.063$, $0.938-1$,
{\em Yellow}: $0.063-0.188$, {\em Yellow-Green}: $0.188-0.313$, {\em
Green-Teal}: $0.313-0.438$, {\em Teal}: $0.438-0.563$, {\em Teal-Blue}:
$0.563-0.688$, {\em Purple}: $0.688-0.813$, {\em Pink}: $0.813-0.938$\}. In the
\cite{jern2013probabilistic} experiment, the hue of each experimenter-defined
category was selected from one of six possible values, each of which falls into
one of the color groups above (two color groups were not used as a possible hue
for the experimenter-defined categories). By categorizing the
participant-generated crystals likewise, we can obtain a broad measure of
category contrast by determining the proportion of participant-generated
crystals that fall into the same groups as the experimenter-defined categories:
If contrast influences the hues of the generated categories, we should observe
minimal overlap between in the color groupings.

\begin{figure}
    \begin{center} \inputpgf{figs/}{jk13-huecontrast.pgf}
        \caption{Analysis of data from \cite{jern2013probabilistic}, Experiment
3. Plotted is the number of generated items that share a color group with one of
the experimenter-defined classes. The ``Expected'' data follows a Binomial
distribution with $p=2/8=1/4$, given there were two experimenter-defined
classes, and eight color groups.}
        \label{fig:jk13-huecontrast}
    \end{center}
\end{figure}

These data, shown in Figure \ref{fig:jk13-huecontrast}, reveal a clear pattern:
The majority of participants in each condition ($n = 22$) generated categories
possessing entirely distinct hues; with 0/6 exemplars sharing a hue with the
experimenter-defined categories. These results can be compared to the
predictions of a Binomial model, which proposes that participants generate hues
at random. That is, if hue selection is not systematic, the probability that any
given example will lie in the same color group as an experimenter-defined
category is given by a Binomial distribution with $p = 2/8=1/4$, as there were
two experimenter-defined categories and eight possible color groups. Chi-square
goodness-of-fit tests reveal that the observed distribution in each condition is
highly inconsistent with the hues being chosen at random, all
$\chi^2(6,N=22)>200$, $p<0.001$. Participants tended to generate items that were
perceptually distinct from the categories they had learned, and were less likely
to generate hues possessed by members of the experimenter-defined categories.

\sxlnote{Added this bit on regularity below - I'm a little undecided as to leave
  it in the main text or place as a gigantic footnote? It is a bit of a
  digression, and it's really just here to address Charles' point.... If we want
  to keep this in the main text, I'll have to amend the following paragraph
  ``Re-analyzing the results...'' to improve the transition.}
%
It is possible, however, that this data can be explained by a process that does
not involve contrast. Specifically, since participants were trained on exemplars
that share the same hue within a category, it is plausible that they generate
exemplars that all share a common hue. If this is the case, given a $6/8 = .75$
probability that a hue distinct from the experimenter-defined categories is
selected, there is a corresponding $.75$ probability (or about 17 out of 22
participants) that the generated category will have no exemplars that share a
hue with any of the experimenter-defined categories. In addition, that there is
a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
category will have all six exemplars that share a hue with any of the
experimenter-defined categories. This prediction of regularity across categories
is much closer to the data than the prediction from a Binomial model and does
not require any mechanism of category contrast.

However, upon analysing the consistency of hues in the generated categories, it
is clear that participants do not generally produce categories with entirely
similar hues across all exemplars. Specifically, only .55 of all generated
categories comprised exemplars with a common hue. Consequently, the assumption
of regularity across categories may be inappropriate for this data set, leaving
the contrast-based explanation as the most plausible for our purposes.

Re-analyzing the results from \cite{jern2013probabilistic} provides preliminary
support that contrast plays a role in category generation. Taken alongside the
analyses reported by \cite{jern2013probabilistic}, our analysis suggests that
generated categories tend to be distinct from {\em and} distributionally similar
to what is already known. However, it is worth noting that our analysis is still
limited: The color groups defined above are imprecise, and it is not clear that
our color grouping is consistent with the psychological color boundaries
perceived by participants. While we did obtain similar results using a variety
of alternative groupings, the hue dimension used in the
\cite{jern2013probabilistic} study does not lend itself straightforwardly to the
the computation of similarities, and thus we cannot be certain of whether our
coding accurately approximates the psychological space of the stimuli. This
precludes traditional applications of categorization models to their data as it
is usually necessary to encode objects in psychological space in order to
accurately determine the similarity between objects. By consequence, although
these results likely indicate that contrast exerts {\em some} influence, they do
not precisely describe the nature of that influence. In the sections below, we
propose a quantitative framework specifying the role of category contrast in
creative generation.

\subsection{Contrast as exemplar dissimilarity: The PACKER Model}

\label{section:PACKER-definition}

As noted above, the constraint that new concepts should differ from what is
already known has been largely overlooked in previous work. This is no doubt in
part due to the vague definition of what it means for a concept to be
``different'': A generated category may be different from what is already known
in any number of respects. Towards providing a more precise definition of the
role of contrast in creative generation, we formalized contrast in a novel
exemplar model, PACKER ({\em Producing Alike and Contrasting Knowledge using
Exemplar Representations}). PACKER explains category generation as a balance
between two fundamental constraints: The category to be generated should not be
similar to known categories, and exemplars within each category should be
similar to one another. These ideas are implemented within the well-studied
exemplar framework -- the PACKER model is an extension of the influential
Generalized Context Model of categorization
\citep[GCM;][]{nosofsky1984choice,nosofsky1986attention}.

Although, as an exemplar model, one of PACKER's proposals is people represent
categories in terms of a collection of stored exemplars, it is worth emphasizing
that the underlying nature of human category representations is not the primary
concern of our work. The choice to develop PACKER within an exemplar framework
reflects the facts that exemplar models have been throughly evaluated, are
strongly theoretically motivated, and dominate much of the theoretical and
empirical work in categorization. The focus of our work with PACKER concerns the
dual constraints of within- and between-class similarity; it is not difficult to
imagine how such constraints may be instantiated using alternative frameworks
(e.g., \citealp{kurtz2007divergent,love2004sustain,smith2000thirty}; for a
review of categorization models, see \citealp{pothoswills2011}).

Both PACKER and the GCM simulate categorization under the assumption that
learners represent categories as a collection of exemplars, corresponding to the
labeled stimuli they have observed. The exemplars are encoded within a
$k$-dimensional psychological space, and model performance is based on the
amount of similarity between the item to be categorized and the stored
exemplars. Similarity between two examples, $s\left(x_i, x_j\right)$, is
computed as an inverse exponential function of distance
\citep[following][]{attneave1950,shepard1957stimulus,shepard1987toward}:

\begin{equation} s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k
\left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
\label{eq:similarity}
\end{equation}
% 
where $w_k$ is the attention weighting of dimension $k$ ($w_k \geq 0$ and
$\sum_k{w_k} = 1$), accounting for the relative importance of each dimension in
similarity calculations, and $c$ ($c>0$) is a specificity parameter controlling
the spread of exemplar generalization. For simplicity, attention will be
distributed uniformly in our simulations (unless otherwise noted). The value of
$r$ depends on the nature of the experimental conditions being simulated: $r=1$
is appropriate for separable dimensions, whereas $r=2$ is appropriate for
integral dimensions \citep[e.g.,][]{shepard1964attention,garner1974processing}.
In our simulations, we set $r=1$ due to the separable nature of the stimulus
dimensions used in our experiments (see Figure \ref{fig:e1-conditions}).

PACKER (as well as its name) was in part inspired by earlier work from the
categorization literature. Specifically, \cite{hidaka2011packing} argued that
natural categories ``pack'' the values of features such that different
categories fill the domain space with distance between one another, while
keeping items within the same category close together. Inspired by this idea,
PACKER proposes that generation is constrained by both similarity to members of
the target category (the category in which a stimulus is being generated) as
well as similarity to members of other categories: the most desirable generation
candidates are similar to members of the target category and not similar to
members of contrast categories. This is achieved by aggregating similarity
across known exemplars differently according to class membership. The aggregated
similarity $a(y,x)$ between generation candidate $y$ and stored exemplars $x$ is
given by:

\begin{equation} a(y, x) = \sum_j{f(x_j) s(y, x_j)}
\end{equation}
% 
where $f(x_j)$ is a function specifying each exemplar's contribution to
generation. A negative value for $f(x_j)$ produces a `repelling' effect (items
are less likely to be generated nearby $x_j$), and a positive value produces an
`attracting' effect (items are more likely to be generated nearby $x_j$). When
$f(x_j)=0$, the exemplar does not contribute to generation.

PACKER sets $f(x_j)$ depending on exemplar $x_j$'s category membership:
$f(x_j) = \theta_t$ if $x_j$ is a member of the target category, and
$f(x_j) = -\theta_c$ if $x_j$ is a member of a contrast category.
$\theta_t$ and $\theta_c$ are free parameters
($0 \leq (\theta_t,\theta_c) \leq 1$) controlling the trade-off
between within- and between-category similarity. For example, when
$\theta_t = \theta_c = 0.5$, $f(x_j) = 0.5$ for members of the
target category and $f(x_j) = -0.5$ for members of other categories; thus, the
model is likely to generate items that are similar to members of the target
category but are not similar to members of other categories. In this way,
$\theta_t > 0$ with $\theta_c = 0$ produces exclusive consideration
of target-category members, and $\theta_c > 0$ with $\theta_t = 0$
produces exclusive consideration of contrast-category members. The combination
of $\theta_t$ and $\theta_c$ parameters thus specifies a wide
breadth of possible approaches; by fitting it to a dataset, one can describe the
relative roles of between-category contrast and within-category similarity in
generation. See Figure \ref{fig:packer-examples} for an illustration of how
these parameters control the relative influence of within-category similarity
and contrast to other categories when generating a new exemplar.

\begin{figure}
	\begin{center} \inputpgf{figs/}{packer-examples.pgf}
      \caption{PACKER generation of a category `B' example, following exposure
        to one member of category `A' and one member of category `B'.
        Predictions are shown for three different parameterizations (differing
        in values of $\theta_t$ and $\theta_c$): {\em (a)}
        Predictions based on contrast similarity only. {\em (b)}
        Predictions based on target similarity only. {\em (c)} Predictions with
        both constraints considered.}
		\label{fig:packer-examples}
	\end{center}
\end{figure}

The probability that a candidate $y$ will be generated is evaluated using an
Exponentiated \citet{luce1977choice} choice rule. Candidates with greater values
of $a(y,x)$ are more likely to be generated than candidates with smaller values:
% 
\begin{equation} p(y \mid x) = \dfrac { \exp \left \{ a \left( y, x
\right) \right \} } { \sum_i{ \exp \left \{ a \left( y_i, x \right)
\right\} } }
\label{eq:packer-choice}
\end{equation}
% 
%where $\theta$ ($\theta \geq 0$) is a free parameter controlling response
%determinism.

It is worth noting that PACKER is only one possible exemplar-based account of
category generation within our proposed framework. That is, PACKER places
specific constraints on the possible values of $f(x_j)$, but other
exemplar-based category generation models with drastically different behavior
can be formalized in this framework by imposing alternative constraints. For
example, as will be discussed in more detail below, PACKER is formally
equivalent to the copy-and-tweak model proposed by \cite{jern2013probabilistic}
when $\theta_c = 0$ and $\theta_t > 0$. Likewise, when
$\theta_t = 0$ and $\theta_c > 0$, PACKER can represent a
contrast-only generation mode, relying exclusively on contrast when generating
new categories. Finally, when $f(x_j) < 0$ for all $x_j$ (regardless of class
membership), a ``pure-packing'' approach is yielded, generating items in
unoccupied areas of the domain. Thus, the proposed framework may be used to
describe a wide variety of qualitatively distinct generation strategies.

It is also worth noting that PACKER represents just one of many possible models
that incorporate contrast in category generation. For example, it may be
possible to extend \cite{jern2013probabilistic}'s hierarchical Bayesian model to
include contrast, but this mechanism would be at odds with the standard
hierarchical Bayesian framework. In contrast, these ideas emerge naturally from
the exemplar view: We have not modified any of the core elements of the GCM in
defining the PACKER model, we simply weight exemplar similarity during
aggregation. Thus, beyond formally specifying the role of contrast in creative
generation, the PACKER account allows us to evaluate the well-understood
principles of the exemplar view within the comparatively understudied field of
conceptual generation. Nonetheless, we will discuss integrating our approaches
in the General Discussion.

\subsubsection{Relation Between PACKER and Copy-And-Tweak}
\label{section:copytweak-packer} The PACKER model is similar to the
copy-and-tweak model reported by \cite{jern2013probabilistic}: Both models are
exemplar-based, and both models generate new items according to their similarity
to known members of the target class. However, PACKER diverges from the
copy-and-tweak model by including a contrast mechanism, enabling generation
according to dissimilarity to members of opposing categories. By consequence,
copy-and-tweak can be realized as a parameterization of the PACKER model that is
insensitive to category contrast. Specifically, when $\theta_c = 0$ and
$\theta_t > 0$ (see Figure \ref{fig:packer-examples}, panel B),
$f(x_j) = 0$ for $x_j$ belonging to contrast categories; thus, PACKER is not
influenced by these items, and is mathematically equivalent to a copy-and-tweak
approach.

In this paper, we report simulations using this copy-and-tweak model. This model
fits within the exemplar-based category generation framework defined above, and
is a continuous-dimension adaptation of the model tested by
\cite{jern2013probabilistic}. By formalizing a model family where PACKER and
copy-and-tweak are different parameterizations of models within the same
framework, the comparison between PACKER and copy-and-tweak provides a test of
the explanatory value of the contrast mechanism based on exemplar dissimilarity:
The account provided by copy-and-tweak will only equal that of PACKER if the
contrast mechanism does not offer an advantage (i.e., if $\theta_c > 0$
significantly improves model fits). Note that the purpose of the article is to
explore and formally analyze the role of contrast in category generation and
thus, we leave extending PACKER to incorporate distributional factors \citep[as
explored by][]{jern2013probabilistic} for future work.

\subsection{Contrast as representativeness: An extension to Jern and Kemp
(2013)}
\label{section:representativeness-definition}

An alternate conceptualization of category contrast is the idea of
representativeness -- exemplars are generated such that they are more
representative of the target category and thus more distinct from and less
similar to other categories. We adopt the definition specified by
\cite{tenenbaum01}, where the representativeness $R$ of an item (or in our case,
an exemplar) $y$ is the relative amount of evidence that is provided by $y$ for
a given hypothesis $h$ compared to all other hypotheses $H'$. More formally,
\begin{equation}
  R(x,h) = \log \dfrac{p(x|h)}{\sum_{h'\in H'}{p(x|h')p(h')}} , 
\label{eq:representativeness}
\end{equation}

In Equation \ref{eq:representativeness}, $p(h')$ is the prior probability
distribution on the hypothesis space that excludes $h$. Note that when there are
only two hypotheses (e.g. two categories) involved in the domain, this prior
takes the value of $1$ since there is ever only one alternative hypothesis.

This Bayesian formalization of representativeness makes it straightforward to
implement into the existing hieriarchical Bayesian model of category generation
developed by \cite{jern2013probabilistic}. Specifically, both models assume that
exemplars are sampled from a given category distribution ($h$ in
Equation \ref{eq:representativeness}). Each category distribution is a
multivariate normal parameterized by a location vector $\mu_k$ and covariance
matrix $\Sigma_k$ (i.e., $h = \mathcal{N}(\mu_k,\Sigma_k)$). The parameters
$\mu_k$ and $\Sigma_k$ are assumed to be samples from a prior
normal-inverse-Wishart (NIW) distribution parameterized by
$\mu_0,\Sigma_0,\kappa,\nu$. Mathematically,
\begin{align}
  \mu_k, \Sigma_k|\mu_0,\Sigma_0,\kappa,\nu &\sim
                                              \textrm{NIW}(\mu_0,\Sigma_0,\kappa,\nu)
                                              \label{eq:mvnsample}\\
  y|C &\sim \mathcal{N}(\mu_k,\Sigma_k) \label{eq:exemplarsample}
\end{align}

In our simulations, we set the prior mean $\mu_0$ to the center of the stimulus
space and $\Sigma_0 = \lambda \mathbf{I}$ where $\lambda$ is a free parameter
and $\mathbf{I}$ is a $d$-by-$d$ identity matrix with $d$ representing the
number of dimensions or features in the domain. The $\kappa$ and $\nu$
parameters are freely estimated within the constraints $\kappa > d-1$ and
$\nu > 0$.

With the assumption of the NIW prior, the location vector $\mu_k$ can be computed
with:
\begin{equation}
  \mu_k = \dfrac
    {\kappa\mu_{0} + n_k \bar{x}_k}
    {\kappa + n_k}
    \label{eq:category_mus}
\end{equation}
where $n_k$ is the number of observed exemplars in category $k$ and $\bar{x}_k$
is the observed category mean. Note that if there are no observed exemplars in
the target category (i.e., if the model is generating a completely novel
category), Equation \ref{eq:category_mus} simplifies to $\mu_k = \mu_0$.

The NIW prior also allows us to infer the category covariance matrix $\Sigma_k$
by computing the following:
\begin{equation}
  \Sigma_k = [\Sigma_D\nu + C_k +
    \dfrac
    {\kappa n_k}
    {\kappa + n_k}
    (\bar{x_k}-\mu_k)(\bar{x_k}-\mu_k)^T
  ] (\nu + n_k)^{-1}
  \label{eq:Sigma_y}
\end{equation}
where $C_k$ is the observed category covariance and $\Sigma_D$ is the domain
covariance matrix from which $\Sigma_k$ samples are obtained. We can
infer $\Sigma_D$ from the observed category covariances $C$ and the prior
covariance $\Sigma_0$. Specifically,
\begin{equation}
    \Sigma_D = \Sigma_0 + \sum_{k}{C_k}
\end{equation}

At this point, both the representativeness model and the hierarchical Bayesian
model from \cite{jern2013probabilistic} are largely identical. However, both
models diverge in their computation of exemplar generation probabilities. While
the original hierarchical Bayesian model produces novel exemplars with
probabilities proportional to the multivariate normal likelihood $p(y|h)$, the
representativeness model generates new exemplars with probabilities weighted by
their representativeness $R(y,h)$. In practice, the probability of generating a
particular candidate $y$ is obtained using an Exponentiated
\citet{luce1977choice} choice rule:
\begin{equation}
  p(y) = \dfrac
    {\exp( \theta \cdot R(y,h))}
    {\sum_k \exp(\theta  \cdot R(y_k,h_k))} .
\label{eq:representative-choice}
\end{equation}
where $\theta$ is a freely estimated response determinism parameter (constrained
such that $\theta \geq 0$).

Despite both the original hierarchical Bayesian model and our representativeness
model sharing identical hierarchical structures, their distinct response
processes can yield very different exemplars. Specifically, the hierarchical
Bayesian model emphasizes the generation of exemplars that maintain
distributional commonalities across categories. In contrast, the
representativeness model focuses on the generation of exemplars that are
representative of the underlying distribution for a target category. Generally,
this representativeness mechanism results in the generation of exemplars that
are less similar (i.e., more different) to contrast categories. We see this
illustrated in Figure \ref{fig:representative-examples}, where the
representativeness model displays a strong preference for generating exemplars
from target category 'B' that are further away from the contrast category 'A'.

\begin{figure}
	\begin{center} \inputpgf{figs/}{hbm-examples.pgf}
		\caption{Generation of a category `B' example, following exposure
to one member of category `A' and one member of category `B'. Predictions are
shown for {\em(a)} the original hierarchical Bayesian model and {\em(b)} the
representativeness model.}
		\label{fig:representative-examples}
	\end{center}
\end{figure}

\subsection{Synopsis and Prognosis}

Research on the creative generation of novel concepts has focused on the finding
that generated categories tend to possess distributional commonalities with
known categories. However, a fundamental goal of concept generation is to create
something {\em new} (i.e., different from what is already known). The manner in
which generated categories differ from known ones is, nonetheless, poorly
understood: Existing theories do not make strong predictions about how
creatively generated concepts should systematically differ from existing ones.
Above, we found encouraging initial support that contrast influences category
generation \citep[][Experiment 3]{jern2013probabilistic}, and we introduced two
novel approaches to formalizing the roles of similarity and contrast in creative
generation.

In the sections below, we present two experiments demonstrating systematic
effects of category contrast on category generation inspired by factors
influencing how the contrast models generate new categories. Our experiments are
based on \cite{jern2013probabilistic}'s paradigm, which is a straightforward
translation of the traditional artificial classification paradigm to the task of
category generation: Participants are first exposed to a single,
experimenter-defined category, and are then asked to generate members of a new
category. We then report formal analyses comparing the exemplar dissimilarity
and the representativeness accounts of our results to that of the hierarchical
Bayesian and copy-and-tweak models developed by \cite{jern2013probabilistic}.
After analyzing the category generation studies, we present third experiment to
demonstrate the effect of contrast in category learning. This ensures that our
results are not idiosyncratic to category generation.

\section{Experiment 1}

To begin our investigation, we sought to extend the early evidence obtained from
our analysis of the \cite{jern2013probabilistic} data, under a variety of
learning conditions and using more standard stimulus materials. We used an
artificial stimulus design: A two dimensional domain of squares, varying in
color and size (see Figure \ref{fig:e1-conditions}, panel A). These dimensions
have been used in numerous classification learning studies
\cite[e.g.,][]{conaway2016similar,conaway2016generalization,shepard1961learning,nosofsky1994comparing}.
Unlike those used in the \cite{jern2013probabilistic} experiments, distance on
these physical dimensions aligns more directly with perceptual similarity,
allowing us to evaluate the role of category contrast in creative generation
more precisely. To extend the evidence provided by the
\cite{jern2013probabilistic} data, we tested the effects of category contrast
after learning one category from a set of qualitatively distinct category
structures, as shown in Figure \ref{fig:e1-conditions}.

Figures \ref{fig:e1-conditions}b-d show the values of exemplar dimensions
belonging to the experimenter-defined categories (`A', or `Alpha') that
participants were assigned to learn about prior to generating a new category.
Each participant learned one of the category types during training. In the
`Cluster' type, category A is a tight cluster of examples in the space.
Perceptually instantiated, the members of category A might, for example, be
large and dark in color. In the `Row' type, category A has a row pattern across
the space, varying along one feature but not the other. Thus, its members might
all be dark in color but would vary in size. Finally, in the `XOR' type, the
experimenter-defined category consists of two clusters separated in opposite
corners of the space, conforming to the exclusive-or logical structure (e.g.,
members are small and dark or large and light).

It should be noted that in our experiments the assignment between the perceptual
and conceptual dimensions (e.g., $X \rightarrow Size$, $Y \rightarrow Color$),
as well as the direction of variation along each dimension (e.g., $dark
\rightarrow light$ or $light \rightarrow dark$) was counterbalanced across
participants. The category types in Figure \ref{fig:e1-conditions} are plotted
in a conceptual space, rather than a perceptual space. Thus, while the
conceptual organization of the category types remains constant, each category
type may have a different physical instantiation according to the counterbalance
assignment. For example, the Cluster type may be large and dark in color, or it
may be small and light in color, depending on the assignment and direction of
the dimensions. For this reason, below we will discuss generation within a
conceptual space, rather than a physically instantiated one.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e1-conditions.pgf}
    \caption{Stimulus domain and category types tested in Experiment 1. Stimuli
are not drawn to scale.}
    \label{fig:e1-conditions}
    \end{center}
\end{figure}

After learning about an experimenter-defined category, participants are asked to
generate examples of a new category. Within this paradigm, an effect of category
contrast would be realized if participants prefer to generate items in locations
that are distant (i.e., perceptually dissimilar) from members of category A.
However, generation is left unconstrained. Critically, participants were not
asked to generate something different in the prompt. For example, participants
assigned to the Cluster condition may generate a tightly clustered category in
the corner opposite of the experimenter-defined category. Alternatively, they
may generate a tightly clustered category directly overlapping with the
experimenter-defined category. Further, they may even generate an entirely
different type of category (e.g., a row category).

Our experimental results also provide a converging test of the classic finding
that generated categories tend to share distributional properties with known
categories in the domain \citep{jern2013probabilistic,ward1994structured}. From
these results, we can predict that, in each condition, participants should
generate categories that are distributionally similar to the
experimenter-defined category: In the Cluster condition, generated categories
should be tightly clustered. In the Row condition, generated categories should
vary more along the X-axis than the Y-axis. In XOR condition, generated
categories should be widely distributed across both dimensions, and the two
dimensions should be positively correlated.

Interestingly, the XOR condition also offers a dissociation between the roles of
category contrast and the emulation of distributional structure:
widely-distributed, positively-correlated categories would need to lie along the
positive diagonal of the space (that is the only place they ``fit''), which is
already occupied by the experimenter-defined category. Thus, if contrast plays a
role, exemplars in the generated categories of participants in the XOR condition
may not be positively correlated -- they may be negatively correlated instead.
In this case, contrast and statistical regularities would interact, which would
be inconsistent with the leading approach in conceptual generation
\citep{jern2013probabilistic}

\subsection{Participants and Materials}

183 participants were recruited from Amazon Mechanical Turk. Each participant
was randomly assigned to one condition: 64 participants were assigned to the
Cluster condition, 61 were assigned to the Row condition, and 58 were assigned
to the XOR condition (sample sizes differ due to random assignment). Stimuli
were squares varying in color (grayscale 9.8\%--90.2\%) and side length
(3.0--5.8cm), see Figure \ref{fig:e1-conditions}. The assignment of perceptual
features (color, size) to axes of the domain space (x, y), as well as the
direction of variation along each axis (e.g., $dark \rightarrow light$ or $light
\rightarrow dark$) was counterbalanced across participants.

\subsection{Procedure}

As noted in the introduction of this paper, the task of generating members of a
new category can be aligned with common tasks studied in the categorization
literature: Whereas classification consists of predicting an object's category
label on the basis of its features, inference consists of predicting an observed
feature, given a set of observed features and a category label. Generation thus
consists of predicting {\em all} features of an object, given a novel category
label. In keeping with this insight, we designed our generation task as an
extension of the traditional artificial classification learning paradigm. The
task differs from traditional work in creative cognition primarily through the
use of an artificial domain, which enables the application of computational
models. The use of an artificial domain also requires the addition of a training
phase, during which participants learn about the categories in the domain. As a
result, unlike most previous studies \citep[e.g.,][]{ward1994structured},
participants in our studies have no experience with the domain before the start
of the experiment, and the experimenter-defined categories are not
hierarchically structured (as are many natural categories).

Participants began the experiment with a short training phase (3 blocks of 4
trials), where they observed exemplars belonging to the `Alpha' category.
Participants were instructed to learn as much as they can about the Alpha
category, and that they would answer a series of test questions afterwards. On
each trial, a single Alpha category exemplar was presented, and participants
were given as much time as they desired to observe it before moving on to the
next trial. Each block consisted of a single presentation of each of the members
of the Alpha category, in a random order. Participants were shown the range of
possible colors and sizes prior to training.

Following the training phase, participants were asked to generate four examples
belonging to another category called `Beta'. As in
\citet{jern2013probabilistic}, generation was completed using a sliding-scale
interface. Two scales controlled the values of the two dimensions (color, size)
for the generated example. An on-screen preview of the example updated whenever
one of the features was changed. Participants could generate any example along
an evenly-spaced 9x9 grid (including members of the Alpha category), except for
any previously generated Beta exemplars. Neither the members of the Alpha
category nor the previously generated Beta examples were visible during
generation. Prior to beginning the generation phase, participants read the
following instructions:

\begin{displayquote} As it turns out, there is another category of geometric
figures called ``Beta''. Instead of showing you examples of the Beta category,
we would like to know what you think is likely to be in the Beta category.

You will now be given the chance to create examples of any size or color in
order to show what you expect about the Beta category. You will be asked to
produce 4 Beta examples - they can be quite similar or quite different to each
other, depending on what you think makes the most sense for the category.

Each example needs to be unique, but the computer will let you know if you
accidentally create a repeat.
\end{displayquote}

\subsection{Results}

We observed a substantial degree of individual differences in our data. In
Figure \ref{fig:e1-samples} we have plotted sample data from several
participants, from which it is evident that different participants generated
qualitatively different category structures. In this section we will focus on
analyzing the data in aggregate, but in later sections we will explore how
individual differences can be explained.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e1-samples.pgf}
    \caption{Sample categories generated by participants in Experiment 1.
Representative samples from common generation profiles are shown.}
    \label{fig:e1-samples}
    \end{center}
\end{figure}

To evaluate the role of contrast, we computed the number of times each stimulus
was generated, as a function of its average city-block distance from members of
the experimenter-defined ``Alpha'' category. These data, shown in Figure
\ref{fig:e1-distanceplots}, reveal a clear pattern: Examples that are more
distant from members of the experimenter-defined categories are more likely to
be generated into a new category. This supports the notion that contrast is a
fundamental constraint on creative cognition and that suggests that statistical
regularity alone is insufficient.

Figure \ref{fig:e1-distanceplots} also depicts, for each participant, the
average distance of members within the generated category ({\em within-category}
distance) against the average distance between members of the generated and
experimenter-defined category ({\em between-category} distance). The narrow
distribution of between-category distances in the XOR condition reflects the
widely distributed nature of the experimenter-defined category, reducing the
possible distances to members of the participant-generated category. These data
reveal a systematic pattern: The majority of participants generated categories
with greater between-category distance than within-category distance. That is,
members of the generated category tended to be more similar to one another than
to members of the experimenter-defined category. To evaluate this claim
quantitatively, we conducted t-tests comparing the amount of within- and
between- class distance in each condition. All conditions possessed greater
between-category distance: Cluster, $t(63) = 11.43$, $p < 0.001$; Row, $t(60) =
13.16$, $p < 0.001$; and XOR, $t(57) = 3.64$, $p < 0.001$. These results provide
further evidence of an effect of category contrast: Participants prefer to
generate categories that are dissimilar to the learned category but maintain
some level of internal cohesion.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e1-distanceplots.pgf}
    \caption{Experiment 1 results. {\em Left}: Frequency of generation as a
function of distance from members of the experimenter-defined category. {\em
Right}: Scatter plot of within-category versus between-category distance in each
of the participant-generated categories.}
    \label{fig:e1-distanceplots}
    \end{center}
\end{figure}

A secondary goal of this experiment was to examine whether we replicate the
classic result that generated categories often possess the same distributional
properties as previously-known categories. For each generated category, we
computed the category range along each axis (X, Y), as well as the correlation
between features. These data, shown in Figure \ref{fig:e1-statsboxes}, reveal
broad individual differences: Within each condition, participants generated
categories spanning the entire X- and Y- axis as well as categories that spanned
very little along each. Likewise, in each condition participants generated
categories possessing strongly positive, neutral, and strongly negative
correlations between the dimensions. Comparing the distributional statistics
between conditions yields a broad yet, as we will see, misleading replication of
the classic effect.

With respect to ranges along each axis (X, Y), the generated categories from
each condition tend to reflect the ranges of the experimenter-defined
categories. The categories generated in the Cluster condition were less widely
distributed along the X-axis compared to Row, $t(123) = 5.61$, $p < 0.001$, and
XOR, $t(120) = 2.68$, $p = 0.008$. Categories generated in the XOR condition
were also less widely distributed along the X-axis compared to Row, $t(117) =
2.56$, $p = 0.012$. This latter effect was not expected because the
experimenter-defined categories for XOR and Row had similar X-ranges. However,
the key finding is that categories from the Cluster condition tended to be more
tightly distributed along the X-axis.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e1-statsboxes.pgf}
    \caption{Box-plots of the distributional statistics from the categories
generated in Experiment 1. Boxes depict the median and quartiles of each
condition, with whiskers placed at 1.5 IQR. All points outside this region are
marked individually.}
    \label{fig:e1-statsboxes}
    \end{center}
\end{figure}

Likewise, categories generated in the Row condition had less Y-axis range
compared to Cluster, $t(123) = 4.57$, $p < 0.001$ and XOR, $t(117) = 9.26$, $p <
0.001$, and categories from the Cluster condition had less Y-axis range compared
to XOR, $t(120) = 3.95$, $p < 0.001$. As expected, the correlations in the
Cluster and Row conditions were not systematically positive or negative ($ps >
0.1$). However, the generated categories in the XOR condition tended to possess
{\em negatively} correlated dimensions, $t(57) = 2.04$, $p = 0.046$. This
finding is notable, as it is the opposite of what would be expected, assuming
learners are emulating the distributional structure of the experimenter-defined
class (which possesses perfectly positively correlated features).

We believe that the failure to replicate this finding is because participants in
\cite{jern2013probabilistic} could differentiate the generated category on a
third dimension (hue) to maintain the statistical regularities on the other two
dimensions. Although the correlation in the XOR condition is significantly
negative, it is clear from the box-plot in Figure~\ref{fig:e1-statsboxes} that
it would be inappropriate to make a strong conclusion (e.g., the median is close
to zero). However, we can conclude with confidence that there are situations
where people do not emulate the distributional structure of the given category.
This indicates that there is more to category generation than the emulation of
distributional structure of other categories in the domain. Further, as we will
discuss in more detail in the model-based analysis section, this is expected by
our proposal that contrast is a fundamental principle in category generation.

\subsection{Discussion} In Experiment 1 we sought to extend our analysis of the
\cite{jern2013probabilistic} data by evaluating the influence of category
contrast on creative generation, given qualitatively different types of prior
knowledge. We found strong evidence for effects of category contrast in each
condition: Participants were more likely to generate stimuli that are more
distant from (i.e., less similar to) members of a previously-learned category,
and members of participant-generated categories tended to be more similar to one
another than to members of previously-learned categories. We also partially
replicated the classic finding that the distributional structure of generated
categories reflects that of previously learned categories
\citep{jern2013probabilistic,ward1994structured}: Members of generated
categories were more widely distributed along dimensions which were widely
distributed in the experimenter-defined category.

Notably, however, we also found that participants who learned an XOR category
(composed of exemplars following a positive diagonal, see Figure
\ref{fig:e1-conditions}) tended to generate items according to a {\em negative}
feature correlation -- the opposite of what was present in the previously
learned category. While this may be difficult to account for under existing
theoretical approaches (which assume generated categories follow the same
distributional structure as known categories), it can be concisely explained
from a category contrast perspective. Specifically, within the XOR condition,
individuals who seek to generate a category that is perceptually distinct from
what is already known are left with only the upper-left and bottom-right
quadrants of the space, as members of the previously-learned XOR category lie in
the bottom-left and top-right. If examples are generated into both of the
available quadrants, the generated category will possess a strongly negative
correlation, opposing that of the experimenter-defined class.

Thus, the core results of Experiment 1 indicate that generated categories can
systematically differ from what we would expect based on prior work. The
negative (or null) correlations observed in the XOR condition suggests an
interesting interaction between contrast with a given category and emulation of
statistical properties. That is, the constraints on creative generation imposed
by category contrast may not simply influence the {\em location} of generated
categories, but also their distributional structure. In Experiment 2, we test
this claim more systematically.

\section{Experiment 2.}

\begin{figure}
    \begin{center} \input{figs/e2-conditions.pgf}
    \caption{Category types tested in Experiment 2.}
    \label{fig:e2-conditions}
    \end{center}
\end{figure}

To test whether category contrast influences the distributional structure of
generated categories, we sought to identify conditions in which differences in
the distributional structure of generated categories cannot be explained by the
distributional structure of the experimenter-defined category. We created two
new category types (depicted in Figure \ref{fig:e2-conditions}) that possess
identical distributional structures (both are tight clusters of examples with no
correlation between features), as they only differ in their Y-axis position: the
`Bottom' category lies in the bottom-center of the space, and the `Middle'
category lies in the center. The distributional equality of these conditions is
key to the design of the experiment: If the distributional structure of
previously learned categories were the only influence on the generated
categories, we should observe no difference in the categories participants
generate between these two conditions. Will participants distribute their
generated category differently between conditions due to the differences in the
available stimulus space for generating a new category?

If category contrast influences the distributional structure of the categories
people generate, then we should observe different types of categories according
to the shape of the space that is {\em unoccupied} by members of previously
learned categories. The difference in the Y-axis position between the Bottom and
Middle conditions produces a considerable change to the shape of the unoccupied
space. Participants assigned to learn the Bottom category should be less likely
to generate exemplars into the lower regions of the stimulus space (as these
areas possess greater similarity to members of the Bottom category), preferring
instead to distribute exemplars across the upper region of the space. This
constraint is lifted in the Middle condition, as the Middle category exemplars
are equidistant to the upper and lower regions of the space. Accordingly,
participants should be more likely to utilize both of these areas. Thus, if
category contrast influences the distributional structure of generated
categories, we should observe more participants in the Middle condition that
generate examples above {\em and} below the experimenter-defined category.

\subsection{Participants, Materials, and Procedure}

122 participants were recruited from Amazon Mechanical Turk. 61 participants
were randomly assigned to the Middle and Bottom conditions each. The stimulus
space and procedure were exactly as in Experiment 1. Participants first
completed a short training phase, followed by the generation phase. The only
difference from Experiment 1 was the category types given to participants.

\subsection{Results}

\begin{figure}
    \begin{center} \inputpgf{figs/}{e2-samples.pgf}
    \caption{Sample categories generated in Experiment 2. }
    \label{fig:e2-samples}
    \end{center}
\end{figure}

As in Experiment 1, we observed broad differences in the generation approach
taken by different participants. To characterize the nature of these
differences, Figure \ref{fig:e2-samples} depicts sample categories generated by
participants. The data from each condition are organized into four columns based
on commonly observed patterns of generation: a `Cluster' type of
tightly-clustered examples, `Row' and `Column' types of exemplars widely
distributed along the one axis but narrowly along the other, and a `Corners'
type, wherein participants placed exemplars in disparate corners of the space.
As before, in this section we focus on analyzing the data in aggregate, but in
later sections we will focus more specifically on explaining the individual
differences.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e2-distanceplots.pgf}
    \caption{Experiment 2 results. {\em Left}: Frequency of generation as a
function of distance from members of the experimenter-defined category. {\em
Right}: Scatter plot of within-category versus between-category distance in each
of the participant-generated categories.}
    \label{fig:e2-distanceplots}
    \end{center}
\end{figure}

We began our analysis by testing for the broad influence of category contrast on
generation. As in Experiment 1, we computed the frequency each stimulus was
generated as a function of its average distance from members of the
experimenter-defined category, as well as each participant's average within- and
between- category distance. These data, shown in Figure
\ref{fig:e2-distanceplots}, yield very similar results. Participants generated
stimuli that are distant from members of the experimenter-defined category, and
the categories in each condition tended to possess more between-category than
within-category distance: Bottom, $t(60) = 5.5$, $p < 0.001$; Middle, $t(60) =
2.71$, $p < 0.01$. We did, however, observe a notable subgroup of participants
in each condition who generated categories with more within-category than
between-category distance. Upon manual inspection, many of these individuals
appear to have assumed a `Corners' strategy, placing exemplars in disparate
corners of the space, thus producing much more within-category distance, see
Figure \ref{fig:e2-samples} for examples.

To explore the distributional structure of the generated categories, we computed
the range of exemplars along each axis (X, Y), as well as the correlation
between features. These data, shown in Figure \ref{fig:e2-statsboxes}, again
demonstrate the degree of individual differences observed in our study. In each
condition, we observed tightly clustered and widely distributed categories along
each dimension, as well as positively, negatively, and uncorrelated categories.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e2-statsboxes.pgf}
    \caption{Box-plots of the distributional statistics from the categories
generated in Experiment 2. }
    \label{fig:e2-statsboxes}
    \end{center}
\end{figure}

As noted above, if the distributional structure of generated categories is
influenced by the shape of the space not occupied by members of known
categories, then participants in the Middle condition would be more likely to
place exemplars in the upper {\em and} lower regions of the space, as members of
the experimenter-defined category are equidistant from these regions.
Participants in the Bottom condition should be less likely to generate category
members in the bottom regions because members of the experimenter-defined
category are located there. One way to test these predictions is to analyze the
Y-axis ranges of the generated categories: If Middle participants utilize the
upper and lower regions of the space, their categories should vary more along
the Y-axis. T-Tests comparing the conditions on the distributional statistics,
however, reveal few between-group differences: the conditions do not differ with
respect to X-axis range, Y-axis range, or feature correlations ($ps>0.17$).

However, our ability to detect differences in Y-Axis range using a standard
$t$-test between the conditions is, in this case, diminished due to the
non-normality of the data (Shapiro-Wilk normality test $W=0.77, p< 0.001$ for
the Middle condition and $W=0.85, p < 0.001$ for the Bottom condition). Figure
\ref{fig:e2-yranges} depicts the Y-axis position of the exemplars generated
within each participant's category. The categories are sorted by overall range,
then by condition assignment. These data reveal that there were nearly as many
participants who generated categories spanning the entire Y-axis as those who
generated categories spanning almost none of the Y-axis. The non-normality of
the Y-axis range distributions thus requires that we use a different approach to
addressing the experiment's main question.

\begin{figure}
    \begin{center} \inputpgf{figs/}{e2-yranges.pgf}
    \caption{Y-Axis range and position of the participant-generated categories
from Experiment 2. Each line corresponds to a participant's category, with
notches corresponding to the Y-axis position of exemplars within the category
(notches may overlap). Participants are sorted by overall range, and then by
condition. }
    \label{fig:e2-yranges}
    \end{center}
\end{figure}

Because our main prediction concerns the generation of exemplars within the
upper and lower regions of the domain, we compared the conditions in terms of
the frequency with which participants generated examples above and below the
categories. Specifically, we counted the number of participants in each
condition who placed at least one `Beta' exemplar on the top and bottom `rows'
of the space (the maximum and minimum possible Y-axis value, respectively). The
resulting contingencies data are shown in Table \ref{table:e2-subset-table}.

Firstly, it should be noted that nearly every participant utilized the top
and/or bottom rows: only $10 / 122$ participants generated their category
entirely within the interior region. Fisher's Exact Tests comparing the
conditions reveal that more Middle participants generated an exemplar in the
bottom row, $p < 0.001$, again demonstrating the role of contrast in guiding
where exemplars are generated. The conditions did not differ in use of the top
of the space, $p = 0.16$, however, more Middle participants placed exemplars in
the top {\em and} bottom rows, $p = 0.038$. The latter effect is of interest
here, as it indicates that the shape of the unoccupied space exerts some
influence on the distributional structure of generated categories: Participants
in the Middle condition were more likely to generate a category spanning the
entire Y-axis. Thus, distributional structure of the generated categories can be
influenced by category contrast alone.

\begin{table}
\begin{center}
\caption{Experiment 2 results.}
\label{table:e2-subset-table} \vskip 0.12in
\begin{tabular}{ l r r} \textbf{Middle} & Used top row & No top row \\ \hline
Used bottom row & 28 & 18 \\ No bottom row & 11 & 4 \\ \\ \textbf{Bottom} & Used
top row & No top row \\ \hline Used bottom row & 16 & 8 \\ No bottom row & 31 &
6 \\
\end{tabular}
\end{center}
\end{table}

\subsection{Discussion}

In Experiment 2, we replicated the core findings from Experiment 1. Stimuli are
more likely to be generated if they are distant from exemplars in other
categories, and most participants generate categories with more between-category
than within-category distance. However, we additionally found that the {\em
position} of a previously learned category (rather than its distributional
structure) influences the types of categories people generate: Participants who
learned the `Middle' type were more likely to generate categories spanning the
entire Y-axis of the space. Participants who learned the `Bottom' type were less
likely to do so as a result of the presence of opposite category exemplars in
the lower regions of the space.

This finding cannot be explained from the perspective that the distributional
structure of previously learned categories is the sole determinant of the
distributional structure of generated categories. However, the observed behavior
is expected from a category contrast perspective: Participants seeking to
generate a perceptually distinct category will be more likely to use areas of
space that are unoccupied by exemplars belonging to previously learned
categories. In the Middle condition, the upper and lower regions of space are
equidistant from members of the experimenter-defined category, whereas in the
Bottom condition, the lower region of the space is closer to members of the
experimenter-defined category. Thus, while Middle participants may form
categories around the use of the equally unoccupied areas, the same is not true
for the Bottom condition.

\section{Model-based Analyses}
\sxlnote{For this section I've added the existence of the representativeness
  model but I think discussion of its performance is still lacking compared to PACKER.}
Experiments 1 and 2 revealed systematic and strong effects of category contrast
on creative generation. In this section, we report the results of simulations
with formal models aimed at explaining our observations. Specifically, we
present simulations from the two novel contrast models: PACKER and the
representativeness model, and compare them to two models that do not incorporate
contrast: the copy-and-tweak model (discussed in Section
\ref{section:copytweak-packer}) and an implementation of the hierarchical
Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
Appendix \ref{ap:hsampling-definition}. For our simulations here, the
copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
parameter constrained to be zero. The comparison of this set of models serves to
highlight the explanatory role of contrast in categorization: If contrast
affords little explanatory advantage, then there should be few differences in
performance between PACKER and copy-and-tweak, or between the representativeness
and hierarchical Bayesian model. The comparison between these models can also
emphasize the necessity of contrast and demonstrate that generation cannot be
explained entirely through the emulation of distributional structure. Each model
has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
are insensitive to the distributional structure of learned categories (relying
only on exemplar similarities), the representativeness and hierarchical Bayesian
model generates categories exclusively on the basis of knowledge of how existing
classes are distributed.

Our approach in this section is to first broadly evaluate and compare the
quality of each model's account to our entire dataset (Experiments 1 and 2
combined), then analyze the ability for each model to explain individual
differences in each experiment, and lastly we describe the strengths and
weakness of each model's account of category generation.

\subsection{Parameter-Fitting}

To obtain a global measure of the quality of each model's account, we fitted the
parameters of each model to our entire dataset (Experiments 1 and 2 combined),
using a hill-climbing algorithm which maximized the log-likelihood of the
model's predictions of the observed responses (1220 responses from 305 total
participants). We fitted three parameters in the PACKER model ($c$, $\theta_t$,
and $\theta_c$; see Section \ref{section:PACKER-definition}), as well as four in
the representativeness model and the hierarchical Bayesian model ($\kappa$,
$\lambda$, $\nu$, and $\theta$; see Section
\ref{section:representativeness-definition} and Appendix
\ref{ap:hsampling-definition} respectively). We fitted only two parameters for
the copy-and-tweak model ($c$, and $\theta_t$), as $\theta_c$ is held constant
($\theta_c = 0$). Attention ($w$, see Equation \ref{eq:similarity}) in PACKER
and copy-and-tweak was set uniformly. Parameters were not allowed to vary
between participants or conditions -- the goal was to obtain the best-fitting
values to our entire dataset.

\begin{table} \centering
\caption{Results of model-fitting to the combined datasets from Experiments 1
and 2. Note that smaller AIC values correspond to better model fits (adjusted
for number of parameters)}
\label{table:global-model-fits}
\begin{tabular}{ l l l l} \\ \textbf{PACKER} & \textbf{Copy \& Tweak} &
\textbf{Representativeness} & \textbf{Hierarchical Bayesian} \\ \hline $AIC =
  9095$ & $AIC = 9842$ & $AIC = 8878$ &  $AIC =
9912$ \\ $L = -4545$ & $L = -4919$ & $L = -4435$ & $L = -4952$ \\ $c = 0.483$ &
                                                                                $c = 3.197$ & $\kappa = 11.232$ &
$\kappa < 0.001$ \\ $\theta_c = 3.16$ & $\theta_c = 0$ (fixed) & $\nu = 2.844$ & $\nu = 5.597$ \\
$\theta_t = 3.494$ & $\theta_t = 2.972$ & $\lambda = 1.917$& $\lambda
                                                                       = 0.055$
  \\ & &$\theta = 5.885 $& $\theta = 3.175$
\\
\end{tabular}
\end{table}

Table \ref{table:global-model-fits} contains the results of this fitting
procedure. Due to the uneven number of fitted parameters among the models, we
compare the model fits using the Akaike Information Criterion
\citep[AIC;][]{akaike1974new}, where smaller values correspond to better fits
(discounted by model complexity as measured by the number of parameters). The
same qualitative results were obtained with alternative model comparison metrics
(e.g., BIC, \citealp{schwarz1978estimating}; $AIC_C$,
\citealp{hurvich1989regression}).

Table \ref{table:global-model-fits} contains the AIC values of each model's best
fit, as well as the corresponding log-likelihood ($L$) and the best-fitting
parameter values. These results reveal strong model differentiation: both
contrast models (PACKER and the representativeness model) achieved far better
fits compared to their non-contrast counterparts: copy-and-tweak and the
hierarchical Bayesian model respectively. Interestingly between the contrast
models, the hierarchical model (representativeness) outperformed the
exemplar-based theory (PACKER), whereas between the non-contrast models, the
reverse is observed. Specifically, here the exemplar-based model
(copy-and-tweak) performed somewhat better than the hierarchical Bayesian model.

\sxlnote{The argument in the following two paragraphs appear to be challenged by
  the success of the representativeness model. I might suggest removing them altogether?}
While PACKER's advantage may tentatively be attributed to the model's
sensitivity to category contrast (this will be explored in detail below), the
advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
attributed to its exemplar-based representation, as opposed to the
prototype-based representation assumed by the hierarchical Bayesian model. As
observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
categories we observed were often widely distributed, with no items near the
category prototype. This aspect of the data is inconsistent with the
multivariate normal distributions (similar to prototypes) used to represent
categories in the \cite{jern2013probabilistic} model, but can be easily
accounted for using an exemplar-based approach.

A key distinction between PACKER and the other three models is that only PACKER
is capable of making strong predictions about the location of new category
members when the target class is entirely novel (i.e., no member of the category
has been observed). Under these circumstances, there are no examples to copy,
and thus the copy-and-tweak model predicts that items are generated at random.
Likewise, with no observations on which to condition the category distribution,
the hierarchical Bayesian and representativeness models also picks an item at
random. Crucially, however, the next generated exemplar from the
representativeness model is strongly influenced by contrast as
representativeness. Thus, it is possible that the failure of these models is
simply due to their inability to explain each participant's first trial
(generating the first item in the `Beta' category). To ensure this is not
driving our results, we conducted an identical set of simulations as above,
excluding this trial (leaving 915 responses in the dataset): Again, PACKER
($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
models.

Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
ratio test to compare the two models. PACKER explains the aggregate data
significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
further evidence that category generation is better explained when contrast is
considered.

Through comparison with the copy-and-tweak model, Figure
\ref{fig:packer-loglike} more clearly demonstrates the robustness of the
explanatory gains yielded by PACKER's category contrast mechanism. It displays
the log-likelihood of the participants' results under PACKER as a function of
the $\theta_c$ parameter. The model's other parameters ($c$,
$\theta_t$) were set according to copy-and-tweak's best fits from Table
\ref{table:global-model-fits}, and thus when $\theta_c=0$, the models are
equivalent. The figure clearly shows a ``sweet spot'': a convex region in which
PACKER achieves superior fits as a result of changes to $\theta_c$. The
best fitting values lie well above the value of 0 assumed by the copy-and-tweak
model, which demonstrates the robustness of the contrast effect (though note
PACKER achieves even better fits when its parameters are fitted together, as in
Table \ref{table:global-model-fits}). In sum, the data are better explained when
both within-category similarity and category contrast is considered.

\begin{figure}
    \begin{center} \inputpgf{figs/}{packer-loglike.pgf}
      \caption{PACKER's fit as a function of its prioritization of
        within-category and between-category similarity (using the
        $\theta_c$ parameter) . To facilitate comparison, PACKER's other
        parameters ($c$, $\theta_t$) were set to the best fitting values
        obtained for copy-and-tweak in Table \ref{table:global-model-fits}. }
    \label{fig:packer-loglike}
    \end{center}
\end{figure}

\subsection{Individual Differences}
\label{section:individual-diffs}

As noted in Experiments 1 and 2, we observed a great deal of individual
differences in the types of categories that participants generated. Within each
condition, there were a wide variety of category types, such as row and column
categories (see Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}). The
simulations reported above serve to evaluate the models while considering the
entire dataset, but a secondary goal of any formal account should be to provide
some explanation of how different profiles of performance emerge. Many of the
individual generation profiles we observed can be described within the PACKER
framework, simply by tuning the model's parameters in a principled manner. In
this section, we describe more specifically how the most frequently observed
profiles can be realized.

By manual inspection, it is evident that the most common profiles of generation
consist of: (A) a tightly-distributed `cluster' of examples, (B) `row'- and
`column'-like arrangements (varying widely along one dimension but not the
other), and (C) a `corners' arrangement with examples placed into disparate
corners of the space. These four profiles are distinct in terms of the
distribution of the generated category along each dimension: Whereas the cluster
profile is tightly distributed along both dimensions, the row and column
profiles are tightly distributed along just one dimension. Finally, the corners
profile is widely distributed along both dimensions.

In the framework proposed by PACKER, the cluster and corners profiles arise
based on different prioritization of within-category similarity versus
between-category contrast, and the row and column profiles arise based on the
prioritization of each dimension in the computation of similarity. For example,
in the cluster profile, there is a high degree of within-category similarity
along both dimensions, whereas in the corners profile there is minimal
within-category similarity. Thus, PACKER's proposal is that these individual
differences arise as a result of different priorities: While the tight cluster
configuration can be considered PACKER's `default' mode (as it maximizes
within-category similarity), the corners profile can be produced when
between-category contrast is put at a higher priority (i.e., $\theta_c > \theta_t$).

Likewise, in the row and column profiles, there is a large degree of
within-category similarity along one dimension but not the other. These
differences likely arise due to a differential focus on one dimension over
another, and thus they can be produced by changes to PACKER's attention weights,
$w_1$ and $w_2$ (see Equation \ref{eq:similarity}). Traditionally, the attention
weights in exemplar models are thought to reflect the diagnostic value of each
dimension towards classifying the known category members
\citep{nosofsky1984choice,nosofsky1986attention,kruschke1992alcove}, but within
a generation context the weights specify the importance of within- and
between-category similarity along each dimension. For example, if all of
attention is allocated along the X-axis ($w_1=1$ and $w_2=0$), similarity along
the Y-axis no longer influences performance. As a result, PACKER will create
categories that are more widely distributed along the Y-axis, as similarity is
not taken into account along that dimension. As a general principle,
differentially weighting one dimension will result in the generation of
categories that are more widely distributed along the ignored dimension,
conforming to a row- or column-like arrangement. See Figure
\ref{fig:packer-attention} for a depiction of how attention influences PACKER's
performance.

\begin{figure}
    \begin{center} \inputpgf{figs/}{packer-attention-examples.pgf}
    \caption{PACKER generation of a category `B' example, following exposure to
one member of category `A' and one member of category `B'. Predictions are shown
for different attention settings: {\em (a)} Increased weighting of the X-axis.
{\em (b)} Increased weighting of the Y-axis. {\em (c)} Uniform weighting
(identical to Figure \ref{fig:packer-examples}).}
    \label{fig:packer-attention}
    \end{center}
\end{figure}

As in PACKER, changes in the parameter settings of the copy-and-tweak model can
also be used to produce different patterns of generation. Indeed, as
copy-and-tweak is simply a special case of the PACKER model, the attention
weights operate exactly as described above to produce row- and column-like
categories. However, because the model is not influenced by category contrast,
it is biased toward generating tightly clustered categories, as new items are
always most likely to be generated near known examples of the target category.
Thus, the lack of a contrast mechanism prevents the model from explaining why
some individuals widely distribute their categories to the corners of the space.

Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
no mechanism to account for category contrast, and the model is most likely to
generate new items that are similar to known examples of the target category.
Although this precludes an account of why individuals might assume a cluster
versus corners profile, the covariance matrix specifying the model's prior
domain distribution, $\Sigma_0$, can be used to explain the generation of
row-like and column-like categories. This covariance matrix specifies the amount
of variance assumed along each dimension (as well as the dimensional
correlations) across the domain of categories. The covariance matrix for a newly
generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
the distributions of previously learned categories (see Appendix
\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
coded into $\Sigma_0$ to alter the dimensional variance of generated categories.

As noted above, while the copy-and-tweak and hierarchical Bayesian models
possess mechanisms to explain row- and column-like categories, they cannot
easily explain why some individuals widely distribute their generated categories
into disparate corners of the space. This, however, reveals a more general
limitation: According to the copy-and-tweak and hierarchical Bayesian models,
the distributional structure of generated categories is {\em independent} of
their location within the domain. For example, although the copy-and-tweak or
hierarchical Bayesian models can be parameterized to generate row- or
column-like categories, there is no mechanism in place to ensure that what is
generated will be distinct from what is already known. In the next subsection,
we explore this prediction through an analysis of the interdependence between
distributional structure and location in creative generation.

\subsection{Category Location vs. Distributional Structure}
\label{section:individual-diff}
As noted above, while all three models make clear claims about the internal
structure of generated categories, the copy-and-tweak and hierarchical Bayesian
models do not make any claims about how generated categories should differ from
what is already known. However, as we observed in the results of Experiment 2,
the distributional structure of a category is not always independent of its
location within the domain. To demonstrate this point in more depth, we computed
the X- and Y- axis ranges of every participant-generated category. Taking the
difference between these values ($X-Y$) produces a measure of each category's
orientation in the space: positive difference scores correspond to categories
with more X-axis range (horizontally aligned, `Row' categories), whereas
negative difference scores indicate the opposite (vertically aligned, `Column'
categories). Neutral differences scores indicate there was an equal amount of X-
and Y-axis range, which can be produced by a number of different category types
(`Clusters', `Corners', etc; see Figures \ref{fig:e1-samples} and
\ref{fig:e2-samples}). By plotting, for each possible stimulus, the difference
scores of categories it was generated within, we can relate the distributional
structure of generated categories to their location within the domain.

However, because many stimuli were infrequently generated (such items near
members of the `Alpha' category), we cannot simply compute the empirical average
of the difference scores, as infrequently generated stimuli would be likely to
show artificially strong differences. Instead, we used a Bayesian analysis to
estimate the mean $\mu_x$ on the assumption that the scores $x$ for each
stimulus are normally distributed with an unknown mean and unknown standard
deviation. The conjugate Normal-Inverse Gamma distribution provides a
straightforward method for this estimation:

\begin{equation} \mu_x = \dfrac { \nu_0 \mu_0 + \sum{x} } { \nu_0 + n }
\label{eq:rangediff-bayes}
\end{equation} \ where $\mu_0$ is the prior mean, $\nu_0$ is a prior scale
parameter (controlling the weighting of the $\mu_0$), and $n$ is the number of
categories in which the stimulus was a member (i.e., the number of scores in
$x$). The default assumption is that there is an equal amount of range along the
X- and Y-axes, and so we set $\mu_0 = 0$. Likewise, to give a moderate amount of
weighting to the prior mean we set $\nu_0 = 1$, though the results are robust to
a range of values. Within this approach, the resulting aggregation is a
trade-off between the number of generations and the strength of the range
difference within each generated category. Infrequently generated stimuli, as
well as those with mixed positive and negative scores, are given neutral
difference scores.

\begin{figure}[p]
    \begin{center} 
      \includegraphics[width=\textwidth]{figs/range-diff-gradients.pdf}
      \caption{Behavioral and simulated range difference gradients. Each panel
        shows, for each stimulus, the dimensional orientation of the categories
        it was generated into: vertically aligned `columns' (orange) versus
        horizontally aligned `rows' (purple).}
      \label{fig:range-diff-gradients}
    \end{center}
\end{figure}

The results of our analysis are shown in Figure \ref{fig:range-diff-gradients}
for the experiment and model results\footnote{Prior to plotting, data was also
processed using a Gaussian filter with $\sigma = 0.8$.}. The left-most column of
Figure \ref{fig:range-diff-gradients} displays the effect of category location
and contrast on the distributional structure of the category generated by
participants. These data reveal strong and consistent patterns across all the
conditions we tested in Experiments 1 and 2: Generated categories are more
tightly distributed along the axis in which they are distinct. For example, in
the `Cluster' condition, exemplars in the bottom-left of the space are more
often generated into vertically aligned categories, and exemplars in the
top-right are more often generated into horizontally aligned categories.
Similarly, in the `Bottom' and `Middle' conditions, horizontally aligned
categories are generated above and below the experimenter-defined categories,
while vertically-aligned categories are generated to the sides. In the `Row'
condition, most categories are horizontally aligned, and lie along the upper
areas of the space. There are no strong range difference patterns in the XOR
condition.

These patterns of performance clearly depict the interdependence between the
distributional structure and location of generated concepts. Our results can be
interpreted in terms of local minimization of between-category similarity: By
distributing the generated category away from members of the
experimenter-defined category, participants may increase the degree of
between-category distance without drastically altering the degree of
within-category similarity.

To explore how well the PACKER, copy-and-tweak, representativeness, and
hierarchical Bayesian models explain our findings, we conducted simulations
using an individual-differences approach. As noted in Section
\ref{section:individual-diffs}, row- and column-like categories can be produced
by each model through changes to the weighting of each dimension. Given this
information, we may use the models to simulate each participant's generation
separately, with the importance of each dimension set according to the relative
range of the participant's generated category along each dimension.

In the PACKER and copy-and-tweak models, the attention weights, $w$, specify the
importance of each dimension in the computation of similarity. While there exist
methods to find the optimal attention weighting scheme given a classification
\citep[see][]{vanpaemel2012using}, for simplicity we may assume that the Alpha
and Beta categories are distinct along dimensions that the Betas do not vary on.
Thus, the weighting for a given participant can be computed as:

\begin{equation} w_k = \dfrac {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } {
\sum_k {\exp{ \{ -\theta_w \cdot\text{range}(k)} \} } }
\label{eq:range-weight}
\end{equation}
% 
where $\theta_w$ is a free parameter controlling how differences in range
correspond to differences in weights (functioning similarly to the $\theta$
parameter in each of the models), and $\text{range}(k)$ is the range of examples
generated by the participant along dimension $k$. We used $\theta_w = 1.5$ in
our simulations, though the results are robust and similar for other $\theta_w$
values. The resulting $w$ values are thus inversely proportional to the range of
generated categories along each dimension, with less range corresponding to
greater weighting.

Unlike the PACKER and copy-and-tweak models, the representativeness and
hierarchical Bayesian model's dimensional variances correspond to the assumed
variance of generated categories along each dimension (rather than the inverse
of the variance). Thus, a different transformation is appropriate for
incorporating the weights computed in Equation \ref{eq:range-weight}. For the
hierarchical Bayesian model, we computed the dimensional variances according to:
$\lambda (1-w_k) 2$, where $\lambda$ is a free parameter specifying the overall
assumed variance of the domain, and $2$ corresponds to the number of dimensions
in our experiments\footnote{This calculation applies only in two-dimensional
  domains, where $w_2 = 1-w_1$.}. Under this approach, evenly distributed
weights correspond to an assumed variance of $\lambda$. Likewise, larger values
of $w$, which are produced when the generated category is tightly distributed
along one dimension, correspond to smaller assumed variances.

Each model was used to simulate each participant's generation independently,
with the importance of each dimension set according to the participant's
generated category. The other free parameters within each model were set as in
Table \ref{table:global-model-fits}. Every participant's generation was
simulated 2,000 times; given the 305 participants tested across the two
experiments, each model generated 610,000 categories in total. For comparison
with our behavioral results, we then computed the range difference gradient
identically as with the behavioral data. The results are shown in Figure
\ref{fig:range-diff-gradients}.

As in the more traditional model evaluation analysis described above, the
contrast models (i.e., PACKER and the representativeness model) provided a much
closer match to our behavioral results than the copy-and-tweak and hierarchical
Bayesian models. In all conditions, the contrast models distribute categories
similarly to the behavioral data: Horizontally-aligned categories tend to be
placed above and below members of the experimenter-defined category, and
vertically-aligned categories tend to be placed to the sides. Conversely,
because the copy-and-tweak and hierarchical Bayesian models are insensitive to
category contrast, these models do not produce any systematic patterns of
association between category location and distributional structure. The sole
exception is within the `Row' condition of Experiment 1, in which the majority
of participants generated a `Row'-like category, widely distributed along the
X-axis but not the Y-axis. In these cases, both models are initialized with
weights that produce Row categories, but because category contrast is not
considered, categories are uniformly generated across the entire domain, rather
than concentrated within the upper-regions as observed behaviorally.

\include{methods_results_e3}

\section{General Discussion}

The creative generation of concepts is an immensely intriguing topic, but it is
also complex and difficult to study in a traditional laboratory setting.
Creative generation is most typically studied using a creative cognition
approach \citep{finke1992creative,smith1995creative}, where creative products
\citep[such as drawings of alien plants and animals;][]{ward1994structured} are
analyzed to obtain insight into the processes and representations involved.
While this approach has yielded a great deal of information about the nature of
creative generation, the commonly employed response modalities (e.g., hand-drawn
pictures) make it difficult to develop computational models capable of formally
evaluating the proposed representations and processes. However, the generation
of novel concepts is not altogether different from other behaviors studied in
cognitive psychology: In particular, it can be viewed as an form of inference
using one's knowledge about existing categories
\citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
this paper we investigated the generation of new categories using the empirical
and computational toolkit from the field of categorization.

While the bulk of prior research on the topic has focused on the classic finding
that generated concepts tend to be distributionally similar to known concepts,
there has been little work addressing the role of contrast in creative
generation: How is it that people are able to create something {\em different}
from what is already known? We developed a novel, exemplar-based model, PACKER,
which formally specifies the role of contrast in generation. The model proposes
that categories are represented as exemplars in a multidimensional psychological
space, and generation is constrained both by within-category and
between-category similarity: Exemplars belonging to the same category should be
similar to one another, and exemplars belonging to different categories should
not be similar to one another.

In addition to an analysis of published data \citep[][Experiment
3]{jern2013probabilistic}, we reported two experiments demonstrating systematic
effects of category contrast in creative generation. Members of
participant-generated categories tended to be highly dissimilar from members of
previously-learned categories, and were usually more similar to one another than
to members of other categories. We also observed broad interdependence between
the distributional structure (feature variance, correlation) and physical
instantiation (location within the stimulus space) of generated categories: In
Experiment 2, we found that the unoccupied regions of the domain influenced the
distributional structure of categories, and in both experiments we observed that
participants distributed their generated categories to increase contrast with
what was already known.

We conducted simulations comparing PACKER's account of our results to the other
major proposals for category generation: a ``copy-and-tweak'' model (realized as
a variant of PACKER with no sensitivity to category contrast), and a
hierarchical Bayesian model designed to explain the classic distributional
similarity effect. In all simulations, we found that PACKER's sensitivity to
contrast captured a previously unexplained and unexplored aspect of human
category generation. Further, by measuring PACKER's fit as a function of its
prioritization of within- and between-category similarity, we observed that
considering either constraint exclusively results in a relatively low-quality
account. Instead, PACKER's best results were obtained when both constraints are
considered, indicating that human learners do not generate novel concepts
exclusively on the basis of within-category similarity or between
class-contrast. This finding mirrors our behavioral results and demonstrates
that both constraints influence creative generation.

\subsection{Similarity and Contrast in Cognition}

The proposal that category contrast is a primary constraint in creative
generation is, in some ways, entirely commonsense. To successful create
something \textit{new}, it must be different from what is known. Beyond its role
in creative generation, category contrast is also of fundamental importance in
categorization more broadly. All other factors held constant, new categories are
easier to learn if they are dissimilar to members of other categories, and
knowledge of highly distinct categories is applied more accurately than that of
ill-defined categories \citep{ashby1994categorization,imai1965discriminability}.
Likewise, basic-level categories \citep{rosch1976basic} are thought to be
abstracted in order to maximize within-category similarity while minimizing
between-category similarity. Finally, the act of forming category
representations affects similarity judgments about category members and
nonmembers, with category members being viewed as more similar to one another
than members of other categories
\citep{goldstone1994influences,goldstone2001altering,goldstone1996isolated}.

Beyond categorization, one can find instances of the trade-off between within
and between-class similarity in linguistic categories over perceptual
dimensions. For example, \cite{regier2007} showed that the partitioning of color
categories reflects such a trade-off in a psychological space -- colors are
partitioned into groups with members that are viewed as highly similar to one
another yet distinct from other colors. A similar trade-off can be observed in
phoneme categories. Different exemplars of the same phoneme must be similar to
one another, while contrasting from other phonemes, such that a listener can
infer the appropriate phoneme. This pattern has been found and modeled in the
natural acoustics of American English vowels
\citep{feldman2013,hillenbrand1995}. As linguistic categories must have been
created at some point in human history, it is revealing that the constraints of
emulating distributional structure across categories and having categories
contrast from one another still bias human category generation today.

The dual forces of within-class similarity and between-class contrast influence
cognitive functions in a wide variety of domains. The PACKER model is notable in
that it successfully interprets this trade-off within the domain of creative
generation, and allows us to begin to understand the relatively understudied
processes involved in creative generation through our more well-developed
knowledge of human categorization.

\subsection{Implications for Creative Cognition}

Although the focus of this article has been to address the processes involved in
creative generation using the empirical and quantitative toolkit of traditional
categorization research, our findings and approach have relevant implications
for research in creative cognition. A central focus of the creative cognition
approach has been to explain acts of creativity in terms of the mental
representations and processes that are commonly studied in cognitive psychology
and cognitive science \citep{finke1992creative,smith1995creative}. However,
unlike other fields in the study of cognition, creative cognition research
rarely employs quantitative models to evaluate the explanatory value of such
representations and processes. Our modeling results provide a concrete example
of how formal approaches may be used to gain insight into the nature of creative
cognition.

In addition to demonstrating the utility of formal modeling in the study
creative cognition, the PACKER model more specifically offers an additional
interpretation of some of the field's most central findings. For example,
perhaps the most foundational principle from this literature concerns the
limiting influence of prior knowledge: Individuals create new categories
composed of features from existing classes, and what is created can be
influenced drastically through the introduction of cues or examples
\citep{marsh1999inadvertent,smith1993constraining}. In this paper, we have
identified another important aspect of the constraining influence of prior
knowledge: What is generated cannot be the same as what is already known.
Further, there is systematicity in how generated categories differ from prior
knowledge. The results of our simulations with PACKER demonstrate that this
influence is concisely explained in terms of a trade-off between within-category
similarity and between-category dissimilarity.

Similarly, PACKER may offer an additional interpretation of existing accounts of
creative generation. Most notably, a leading account within the creative
cognition literature, the Path of Least Resistance
\citep{ward1994structured,ward1995s}, also explains generation in terms of an
exemplar-based retrieval process. This account was designed to explain the
creative generation of natural categories (e.g., new species of plants and
animals) and as a result relies strongly on the hierarchical organization of
these categories: Individuals are thought to retrieve an example of the
higher-level category being generated (e.g., \textit{bird} may be retrieved from
the category \textit{animal}), and then systematically alter what was retrieved
to make something new. As the PACKER model does not assume knowledge is
hierarchically organized \citep[this is true of the exemplar view more broadly,
see][]{murphy2016exemplar}, the model may be viewed as a formal instantiation of
the Path of Least Resistance for application in a traditional artificial
categorization domain (when there is no established hierarchy of categories).
PACKER's success in explaining generation within an artificial domain motivates
future work exploring the nature of category contrast within a more naturalistic
setting.

\subsection{Implications For Categorization}

Although our work has mainly addressed the processes involved in the creative
generation of concepts, in this paper we have studied these processes as they
apply within an artificial categorization domain. By consequence, our findings
also provide an advance in our understanding categorization more broadly.

Categorization research addresses the representations and processes that
underlie the learning and use of categories. Category learning tasks are
generally about figuring out which items belong to which category. Once learned,
categories are generally used to classify new stimuli and to make inferences
beyond the available information. Our work is fairly unique in that people learn
a single category through positive examples and then figure out another category
that would make sense in the domain. In other words, only one category is
provided to be learned and the use that is required of that category is to
generate a contrast category. Such propagating of categories (producing one from
another) may not seem like a fundamental or ubiquitous cognitive activity, but
consider navigating a highly unfamiliar environment -- it would be practical,
perhaps even critical, to generate expectations about what other kinds of things
you are likely to encounter based on those that you already have. In the present
studies, we have learned something about the form that such expectations are
likely to take.

We can think of the category generation task in our studies as asking a person
to formulate an idea about what set of items in the domain are most
interestingly {\em not} members of the original category. To meet this
condition, the items must take some form of coherence that aligns with that of
the original category and some form of distinctiveness relative to the original
category. Reflecting the basic level of organization in natural categories, it
makes sense to generate a set of items that possess strong within-category
coherence (by importing or systematically transforming the internal structure of
the original category) as well as strong between-category differentiation (by
creating maximum contrast with the original category, be it through
exemplar-based dissimilarity or the maximization of exemplar-category
representativeness). In this sense one can see the patterns of performance in
the category generation task as recapitulating the order of semantic
organization.

\subsection{Exemplar dissimilarity and Representativeness in Categorization}

\sxlnote{I'm struggling a little with this section! That said, I think this may
  also be a good point to address Charles' concern re: the possible conditions
  under which contrast are important.} 
Although both PACKER and the
representativeness model perform better than the other models, they appear to be
capturing fundamentally different aspects of contrast in categorization.
Specifically, while the representativeness model excelled in predicting our
category generation data, PACKER was the best-fitting model in accounting for
participant error in category learning. The source of this discrepancy is still
not yet clear -- it could be an indication of a fundamental difference between
the processes involved in generation versus learning where exemplar
dissimilarity is a larger factor in category generation while representativeness
exerts more influence in category learning.

Xian: Please
write about the two hypotheses and that followup work needs to look into why the
Bayesian representativeness model doesn't do as well at explaining category
learning ease, but does better at category generation.

MAKE SURE WE CITE AND DESCRIBE \citet{thomas98}! \sxlnote{I moved the
  description and citation of thomas98 to the introduction for justification of
  why we think categories should share similar distributions.}
\subsection{Limitations and Future Directions} Although we have demonstrated
that contrast plays a role in categorization, our results were equivocal as to
whether contrast is governed by exemplar dissimilarity or representativeness.

Although successful in explaining our results in all three studies (albeit not
as well as the representativeness model for Experiments 1 and 2), PACKER does
not provide a full account of what is known about category generation. Most
notably, in this paper we have not evaluated the model's ability to explain the
classic finding that generated categories tend to share distributional
commonalities with previously learned categories
\citep[see][]{jern2013probabilistic,ward1994structured}. While we successfully
replicated this effect in Experiment 1, we also found that its influence was
limited in comparison to the fundamental constraints imposed by category
contrast. Even within Experiment 1, we found systematic inconsistencies: by
generating exemplars into unoccupied regions of the space, participants who
learned an `XOR' category, composed of members that are widely distributed along
both features and are positively correlated in space, tended to generate
categories with an opposite (negative) correlation. More generally, PACKER
inherits the strengths and weaknesses of exemplar models of categorization: It
provides a simple and flexible model that explains many results, but deviates
systematically from human performance in some cases.

Nonetheless, these classic effects are a core element of the phenomenology of
creative generation, and PACKER does not include any mechanisms that explain
them. Instead, through the development and evaluation of the PACKER model, we
have sought to add new elements into such a phenomenology: The broad and strong
influence of category contrast, and the interdependence between category
location and distributional structure. It may be possible to combine the
hierarchical Bayesian approach proposed by \cite{jern2013probabilistic} with
PACKER's underlying claims to obtain a ``best of both worlds'' model, capable of
explaining the role of contrast in creative generation, as well as the emulation
of distributional structure. However, as noted in the introduction, the
incorporation of category contrast is antithetical to the core principles of a
traditional, semi-conjugate Bayesian approach. This suggests that category
generation is a fundamentally different computational-level problem
\citep[different from those posed by][]{jern2013probabilistic,kemp2014taxonomy}.

Characterizing that problem and conducting a rational analysis is an important
direction for future research. To that aim, we plan to explore the connection
between exemplar modeling as an Importance Sampling approximation
\citep{shi10exemplar}, and see what sort of computational-level problem PACKER
approximates. Once formalized in probabilistic terms, it should also be
straightforward to incorporate distributional factors into the model. This would
unite PACKER with the Bayesian representativeness model, and they would differ
in terms of their assumptions about how people represent category distributions
(as exemplars or prototypes, respectively). Alternatively, it may be possible to
integrate the core principles of contrast into other categorization models
\citep[e.g.,][]{love2004sustain,kurtz2007divergent,smith2000thirty}.

Finally, although an overall goal of this work is to advance the study of
creative thought by explaining generation through empirical and quantitative
approaches from research in categorization, these approaches are not easily
aligned with many of the alternative theoretical viewpoints in the broader study
of creativity \citep[for a review see][]{kozbelt2010theories}, such as those
based on free association \citep{mednick1962associative} and conceptual
combination \citep{estes2002emergence,murphy1988comprehending}. Instead, we
reduced our focus by studying a highly complex behavior (creative generation) as
it applies within a well-established domain (artificial category learning). We
hope that future work incorporates and highlights the importance of contrast
into theories of creativity.

\section{Conclusions}

The generation of new concepts and ideas is a highly interesting topic, but it
is difficult to study in a controlled experimental environment. In this paper,
we have provided such an examination of category generation as it applies within
an artificial categorization experiment. Extending the literature on creative
cognition, our experiments provide a detailed picture of the role of category
contrast in generation: People seek to create concepts that are distinct from
what they already know, and the nature of what is created can be influenced by
what does not yet exist. Our simulations with traditional exemplar models, as
well as a hierarchical Bayesian model, provide strong support for the claim that
category contrast is of fundamental importance. More generally, our results
demonstrate that popular explanatory approaches from basic research in cognitive
science can offer a precise, quantitative account of a behavior as complex as
that of creative generation.

\clearpage
\section{Acknowledgments} Previous versions of this work were presented at the
Thirty-Ninth Annual Conference of the Cognitive Science Society and Forty-Ninth
Annual Meeting of the Society for Mathematical Psychology. Support for this
research was provided by the Office of the VCRGE at the UW - Madison with
funding from the WARF. We thank Alan Jern and Charles Kemp for providing code
and data.
\end{flushleft}

% references section
\clearpage \bibliographystyle{apacite} \bibliography{citations.bib}
\clearpage

\appendix \numberwithin{equation}{section}

\section{The Hierarchical Bayesian Model of Concept Generation}
\label{ap:hsampling-definition}

\cite{jern2013probabilistic} demonstrated how a hierarchical Bayesian model
could explain the distributional correspondences between observed and generated
categories. In their model, exemplars of generated categories were viewed as
samples from a multivariate Normal distribution over the dimensions of stimulus
space. The mean of the generated category was independent of the observed
categories, but the covariance matrix (encoding feature variances and
correlations) was based on a common prior distribution. Generating a new
category was thus completed by sampling a new category mean (uniform over
stimulus space) and covariance matrix from the common prior distribution.
Because the shared prior distribution's parameters were unobserved, the
hierarchical Bayesian approach was used to infer its parameters from the
previous categories (their feature variances and correlations), and then to
generate the covariance matrix of the new category.

In our implementation of their model\footnote{Note that
\citet{jern2013probabilistic}'s model is slightly different, as they used a
semi-conjugate model. Their model acts very similarly to our version.}, each
category's exemplars are viewed as samples from a multivariate Normal
distribution with parameters ($\mu, \Sigma$). Category covariance matrices
(specifying variance and covariance along $k$-dimensions), are assumed to be
Normal-Inverse-Wishart distributed with parameters: $\nu$ ($>k-1$), $\kappa$
($>0$), and $\Sigma_D$. $\nu$ and $\kappa$ are treated as free parameters in our
simulations, and $\Sigma_D$ is the domain-wide covariance matrix from which all
categories are viewed as samples. Assuming a given $\Sigma_D$, a category
covariance matrix $\Sigma$ can be computed on the basis of its examples:

\begin{equation} \Sigma = \left[ \Sigma_D \nu + C + \dfrac {\kappa n} {\kappa +
n} (\bar{x}-\mu)(\bar{x}-\mu)^T \right] (\nu + n)^{-1}
\label{eq:Sigma_B}
\end{equation}
% 
where $\bar{x}$ and $C$ are the empirical mean and covariance of the category's
known members, and $n$ is the number of observed members of the category. When
there are fewer than two known members of the category (and thus no covariance
to speak of), $\Sigma = \Sigma_D\nu$.

The category mean, $\mu$, can be computed as:

\begin{equation} \mu = \dfrac {\kappa\mu_{0} + n \bar{x}} {\kappa + n}
    \label{eq:category_mus_jk13}
\end{equation}
% 
where $\mu_{0}$ is the prior mean. In our simulations, $\mu_{0}$ is set to the
center of the domain. However, when no examples of the target category have been
observed, generation is assumed to be random. In practice, the model's best fits
are achieved when the $\kappa$ parameter, which controls the influence of
$\mu_0$ on $\mu$, is set very close to zero (hence, the influence of $\mu_0$ is
minimal).

Importantly, the domain-wide covariance matrix $\Sigma_D$ is unobserved and
needs to be inferred from the observed categories. For conjugacy, if $\Sigma_D$
is viewed as a sample from an Inverse-Wishart distribution with scale
$\Sigma_0$, $\Sigma_D$ can be computed as:

 \begin{equation} \Sigma_D = \Sigma_0 + \sum_y{C_y}
    \label{eq:category_sigmas}
\end{equation}
% 
where $\Sigma_0$ is the prior covariance in the domain. In our simulations,
$\Sigma_0 = \lambda {\bf I}$, where $\lambda$ is a free parameter controlling
the expected variance of dimensions (dimensions of the domain covariance matrix
are expected to be uncorrelated) and ${\bf I}$ is a $k$-by-$k$ identity matrix.

Generated exemplars are drawn from a multivariate Normal distribution specified
by $(\mu, \Sigma)$. Thus, $p(y)$ is
\begin{equation} p(y \mid x) = \dfrac {\exp \left \{ \theta \cdot {\rm
Normal}(y; \mu, \Sigma) \right \} } {\sum_i \exp \left \{ \theta \cdot {\rm
Normal}(y_i; \mu, \Sigma) \right \} }
\end{equation}
% 
where $\theta$ is a response determinism parameter and ${\rm Normal}(y; \mu,
\Sigma)$ denotes a multivariate Normal density evaluated at $y$.

\end{document}