### 2019-07-29 Manuscripts/model-descriptions/negative-weighted/four_models.tex
Updates packer manuscript descriptions (long overdue.) Fixing some representativeness weirdness.

%!TEX output_directory = latex_out/

\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
headheight=15pt]{geometry} \usepackage{amsmath, listings}

\begin{document}

\section{Generating Exemplars from Categories-by-Negation}

When generating exemplars from categories defined by a label that describes what is \emph{not} already learned, we can
define a simple model where these exemplars are samples drawn with a probability proportional to the inverse (or 1-?)category
distribution of

An interesting property of such a model is that it assumes that categories-by-negation are multi-modal, unlike the
unimodal assumptions currently adopted by current models of category generation.