### 2019-09-20 Manuscripts/model-descriptions/negative-weighted/four_models.tex
Added notAvsB subdirectory. Also short writeup on some of the interesting modeling results. Next step, get efficient individual fits working -- treat each ppt as a separate dataset and run 1000 chtc chunks on it.

%!TEX output_directory = latex_out/

\documentclass[12pt]{article} \usepackage[letterpaper, margin=1in,
headheight=15pt]{geometry} \usepackage{amsmath, listings}

\begin{document}

\section{Generating Exemplars from Categories-by-Negation}

When generating exemplars from categories defined by a label that describes what
is \emph{not} already learned, we can imagine a mechanism that focuses on the
areas of the feature space that is not occupied by learned categories. This can
be implemented with a simple model where exemplars from these novel categories
are generated with a probability inversely proportional to the density of all
other categories. For simplicity, we assume that each individual category has
its exemplars distributed as a multivariate Gaussian parameterized by a
$d$-sized vector of empirical means $M$ and a $d$-by-$d$ empirical covariance
matrix $S$, where $d$ describes the dimensionality (i.e., number of features).

More formally, the probability of generating an exemplar $y$ given a set of
categories $C'$\footnote{This model assumes that the novel category is a negated
  category with respect to \emph{all} learned categories. It would be
  appropriate for data from people generating exemplars that are not what they
  previously learned. However, if participants were instead instructed to
  generate exemplars that are not from some subset of a larger set of learned
  categories (e.g., generate not Alphas, having learned both Alpha and Beta
  category exemplars), this simple model would not be appropriate. For our
  experiments so far, we only test participants who have learned one category,
  so there is no problem at this point. For future development, however, we
  might want to consider negative-space models that can consider a subset of
  learned categories. This is probably as simple as having $C'$ represent the
  relevant subset of learned categories.} excluding the novel category can be described as
proportional to a density function $n(y|C')$ where

\begin{equation}\label{eq:negmech}
n(y|C') = -\sum_{c \in C'} \textrm{MvN}(y; M_c,S_c)
\end{equation}

To get the actual choice probabilities, we normalize these values using an
exponentiated Luce's choice rule:

\begin{equation}\label{eq:negluce}
p(y|C') = \dfrac
{\textrm{exp}(\theta \cdot n(y|C'))}
{\sum_i \textrm{exp}(\theta \cdot n(y|C'))}
\end{equation}
where $\theta$ is a response determinism parameter that is freely estimated.

% (An interesting property of such a model is that it assumes that
% categories-by-negation are multi-modal, unlike the unimodal assumptions
% currently adopted by current models of category generation.)

Note that with this simple model the probability of generating an exemplar is
completely dependent on the learned categories -- there is no explicit
within-category constraints defined by the model.

Although this model is limited in its complexity, it is straightforward to apply
as an extension to any of the other models of category generation. We describe
extensions to four models -- PACKER, copy-and-tweak, Jern and Kemp's (2013)
hierarchical Bayesian model, and the representativeness model -- in the following sections.

\section{Extending PACKER and copy-and-tweak}

We begin by briefly describing the structure of the PACKER and copy-and-tweak
models (which can be implemented as a special case of PACKER).

These models predict the generation of novel exemplars by measuring the
candidate exemplar's similarity to other exemplars. Specifically, the similarity
between exemplars $x_i$ and $x_j$ is computed using Shepard's law:

\begin{equation}
s\left(x_i,x_j\right) = \exp \left\{ -c \left[\sum_{k}{ w_k \left| x_{ik} - x_{jk} \right|^r }\right]^{1/r} \right\}
\label{eq:similarity}
\end{equation}
where $k$ indicates each feature, $w_k$ is the weight placed on that feature,
$c$ is a free-to-vary specificity parameer, and $r$ allows us to define the
distance metric. In most of our simulations, we fix $r$ to 1 (implementing a
city-block distance metric). Unless otherwise specified, $w_k$ is fixed to be
equal across all features.

PACKER takes into account the similarity of a candidate exemplar to the existing
novel category (i.e., the target category) exemplars as well as its
dissimilarity to contrast category exemplars. This is implemented as an
aggregated similarity $a$ between candidate $y$ and all other exemplars $x$:

\begin{equation}
  a(y, x) = \sum_j{f(x_j) s(y, x_j)}
\end{equation}
where $f(x_j) = \theta_t$ when $x_j$ is a member of the target category and
$f(x_j) = \theta_c$ when $x_j$ is a member of a contrast category.

To extend this model with the negative-space mechanism, we linearly combine this
aggregated similarity with the densities of the contrast categories $n(.|C')$
(Equation \ref{eq:negmech}):

\begin{equation}\label{eq:negpacker}
b(y,x) = a(y,x) + \gamma \cdot n(y|C')
\end{equation}
where $\gamma$ is a free-to-vary parameter that describes the amount of weight
placed on the negative-space mechanism.

The probability that a given item $y$ will be generated given the model's memory
$x$ is computed using relative summed similarity values across all generation
candidates $y_i$ (i.e., with an exponentiated Luce's choice rule):

\begin{equation}
p(y) = \dfrac
{ \exp  \left \{ b \left( y, x \right) \right \} } 
{ \sum_i{ \exp  \left \{  b \left( y_i, x \right) \right\}  } }
\label{eq:packer-choice}
\end{equation}

The negative-space copy-and-tweak model is instantiated in exactly the same way,
but with $\theta_c$ set to 0, since the model does not account for contrast
category similarity. In addition, $\theta_t$ and $\gamma$ combine to serve as the
response determinism parameter that is free to vary. Response determinism for
PACKER is a combination of $\theta_c$, $\theta_t$, and $\gamma$ parameters.

\section{Extending Hierarchical Bayesian and Representativeness models}

Unlike the exemplar models, the hierarchical Bayesian and Representativeness
models do not explicitly measure exemplar similarity on the feature space.
Instead, both of these Bayesian models assume exemplars $x$ within a category
$c$ are influenced by an underlying multivariate Gaussian distribution
parameterized by means $\mu_c$ and covariance matrix $\Sigma_c$.

$\mu_c$ and $\Sigma_c$ are assumed to be Normal-Inverse-Wishart distributed. The
former can be straightforwardly computed with:
\begin{equation}\label{eq:mu_c}
  \mu_c = \dfrac
    {\kappa\mu_{0} + n_c \bar{x}_c}
    {\kappa + n_c}
\end{equation}
where:
\begin{itemize}
    \setlength\itemsep{-0.5em}
    \item $\mu_{0}$ is the prior mean along $d$ dimensions. Here we set it to the middle of the space.
    \item $\kappa$ is a scalar hyper-parameter, roughly weighting the importance of $\mu_{0}$. $\kappa$ must be greater than zero.
    \item $n_c$ is the number of observations in $x_c$
    \item $\bar{x}_c$ is the sample mean along $d$ dimensions
\end{itemize}

$\Sigma_c$ is influenced by members of the target category $c$ as well as all
other members through a domain-level covariance matrix $\Sigma_D$. Assuming the
observed category covariances $S_c$ are Wishart-distributed, we can compute $\Sigma_D$
with:

\begin{equation}
  \Sigma_D = \Sigma_0 + \sum_{c}{S_c}
\end{equation}
where $\Sigma_0$ is a $d$-by-$d$ prior covariance matrix defined as the
$d$-dimensional identity matrix $I_d$ weighted by a free parameter $lambda$:
\begin{equation}
    \Sigma_0 =  \lambda I_d
\end{equation}

With this, we can compute $\Sigma_c$ as:

\begin{equation}
  \Sigma_c = [\Sigma_D\nu + S_c +
    \dfrac
    {\kappa n_c}
    {\kappa + n_c}
    (\bar{x_c}-\mu_c)(\bar{x_c}-\mu_c)^T
  ] (\nu + n_c)^{-1}
  \label{eq:Sigma_c}
\end{equation}
where $\nu$ is an additional free parameter weighting the importance of
$\Sigma_D$.

As with the other models, choice probabilities are computed using the
exponentiated Luce's choice rule. With some general function $g(\cdot)$:

\begin{equation}\label{eq:hierLuce}
p(y) = \dfrac
{ \exp  \{ \theta \cdot g(\cdot) \} } 
{ \sum_i{ \exp \{ \theta \cdot g(\cdot) \}  } }
\end{equation}

The specific form of $g(\cdot)$ is where the hierarchical Bayesian model
diverges from the representativeness model. For the hierarchical Bayesian model
, $g(\cdot)$ is the density at $y$ under a multivariate Gaussian
parameterized by $\mu_c$ and $\Sigma_c$:

\begin{equation}\label{eq:hierg}
g(\cdot) = \textrm{Normal}(y;\mu_c,\Sigma_c)
\end{equation}

For the representativeness model, $g(\cdot)$ is the representativeness
$R(y,h)$ of $y$ from the target hypothesis $h$ compared to all other hypotheses
$h'$. Here, each hypothesis is represented by the probability density of the
underlying distribution of a specific category. Formally,

\begin{align}
g(\cdot) & = R(y,h)\\
      & = \log \dfrac{p(y|h)}{\sum_{h'\in \mathcal{H}^c}{p(y|h')p(h')}} . 
\label{representativeness}
\end{align}
where $h_c$ is a multivariate Gaussian parameterized by $\mu_c$ and $\Sigma_c$
for a specific category $c$ and $\mathcal{H^c}$ represents the hypothesis set
from all contrast categories.

Similar to the exemplar models, the extension of both hierarchical Bayesian and
the representativeness models are performed by linearly combining the
negative-space densities computed from $n(y|C')$ to the original function prior
to normalization via the exponentiated Luce choice rule. Formally, we can
represent the negative-space extensions of both Bayesian models by amending
Equation \ref{eq:hierLuce} to:

\begin{equation}\label{eq:hierLuceNeg}
p(y) = \dfrac
{ \exp  \{ \theta \cdot g(\cdot) + \gamma \cdot n(y|C') \} } 
{ \sum_i{ \exp \{ \theta \cdot g(\cdot) + \gamma \cdot n(y|C') \}  } }
\end{equation}

Given the structure of the Bayesian models, we adopt a more principled
implementation of $n(y|C')$ by specifying $\mu_c$ (from Equation \ref{eq:mu_c})
and $\Sigma_c$ (from Equation \ref{eq:Sigma_c}) as the means and covariance matrices
of the multivariate Gaussian, instead of the empirical $M_c$ and $S_c$ values.
Specifcally, Equation \ref{eq:negmech} becomes:

\begin{equation}\label{eq:hiernegmech}
n(y|C') = -\sum_{c \in C'} \textrm{MvN}(y; \mu_c,\Sigma_c)
\end{equation}

\end{document}